\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,mathrsfs,bm}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}

\allowdisplaybreaks

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Law}{\mathrm{Law}}
\newcommand{\KL}{\mathrm{KL}}
\newcommand{\Ent}{\mathrm{Ent}}
\newcommand{\TV}{\mathrm{TV}}
\newcommand{\dd}{\,\mathrm{d}}
\newcommand{\mcP}{\mathcal{P}}
\newcommand{\mcL}{\mathcal{L}}
\newcommand{\mcH}{\mathcal{H}}
\newcommand{\mcF}{\mathcal{F}}
\newcommand{\mcK}{\mathcal{K}}
\newcommand{\mcV}{\mathcal{V}}
\newcommand{\one}{\mathbf{1}}

\title{Convergence, Hypocoercivity, and Mean-Field Limit\\
for the Kinetic Core of the Geometric Gas}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
This paper isolates the pure kinetic backbone of the geometric gas: a confining
underdamped Langevin dynamics on $\R^d$, augmented by a row-normalized viscous
velocity alignment force and followed, at observation times, by a smooth
velocity squashing map. The cloning operator, companion sampling, fitness
potential, and state-dependent anisotropic diffusion are deliberately omitted.
Within this kinetic-only setting we prove the two structural facts that make the
model tractable: the viscous force is uniformly bounded on velocity-capped
states, and it is dissipative on relative kinetic energy. We then formulate the
finite-particle hypocoercive entropy estimate that yields an $N$-uniform
logarithmic Sobolev inequality and exponential convergence in relative entropy
and total variation. Finally we derive the McKean--Vlasov limit for the viscous
alignment field, prove quantitative propagation of chaos in relative entropy,
and pass the $N$-uniform logarithmic Sobolev inequality to the mean-field
equilibrium. The paper is self-contained in the sense that all assumptions,
definitions, and theorems used below are stated explicitly in this document.
\end{abstract}

\tableofcontents

\section{Introduction}

The purpose of this paper is to extract and organize the convergence theory of
the kinetic part of the geometric gas in a form that does not depend on the
cloning stage or on the adaptive diffusion geometry. The model studied here is
the following.

\begin{itemize}[leftmargin=1.5em]
\item The position variable evolves by transport: $\dot x = v$.
\item The velocity variable is subject to a confining force $-\nabla U(x)$,
linear friction $-\gamma v$, isotropic Brownian noise, and a row-normalized
velocity alignment force.
\item After each observation interval one may apply a smooth radial squashing
map $\psi$ to cap the observed speed without destroying the previously obtained
probabilistic estimates.
\end{itemize}

Two themes drive the analysis.

\begin{enumerate}[leftmargin=1.5em]
\item The \emph{backbone hypocoercive mechanism} comes from underdamped
Langevin dynamics with confinement and friction.
\item The \emph{new viscous term} is harmless at the level of stability because
its action is bounded and dissipative on relative velocity disagreements.
\end{enumerate}

The paper therefore proceeds in two layers. First we prove the structural
lemmas for the viscous operator and the velocity squashing map. Then we graft
those facts onto the hypocoercive framework for the kinetic backbone,
obtaining $N$-uniform entropy and total-variation convergence. The final
sections pass to the mean-field equation, prove a quantitative propagation of
chaos estimate, and deduce a mean-field logarithmic Sobolev inequality. The
analysis uses only tools from Langevin dynamics, hypocoercivity,
Harris-type ergodic theory, and McKean--Vlasov limits
\cite{roberts1996exponential,meyn2012markov,villani2009hypocoercivity,sznitman1991topics}.

\section{The Kinetic Geometric Gas Core}

\subsection{Standing assumptions}

\begin{assumption}[Strongly confining potential]\label{ass:potential}
The potential $U:\R^d\to\R$ satisfies:
\begin{enumerate}[label=\textnormal{(\roman*)},leftmargin=1.5em]
\item $U\in C^2(\R^d)$ and there exist constants $0<m_U\leq L_U<\infty$ such
that
\[
\|\nabla U(x)-\nabla U(y)\|\leq L_U\|x-y\|
\qquad
\forall x,y\in\R^d.
\]
\item The Hessian is uniformly positive in the sense
\[
\langle \xi,\nabla^2 U(x)\xi\rangle\geq m_U\|\xi\|^2
\qquad
\forall x,\xi\in\R^d.
\]
\item $U$ is coercive in the Lyapunov sense:
\[
\langle x,\nabla U(x)\rangle \geq \alpha_U\|x\|^2-\beta_U
\qquad
\forall x\in\R^d
\]
for some $\alpha_U>0$ and $\beta_U<\infty$.
\end{enumerate}
\end{assumption}

\begin{remark}
Assumption~\ref{ass:potential}(i)--(ii) places the paper in the classical
uniformly convex Langevin regime. In this regime the Gibbs measure
\eqref{eq:gibbs} satisfies a logarithmic Sobolev inequality by the
Bakry--\'Emery criterion, and the modified-entropy computation below yields an
explicit hypocoercive decay rate for the kinetic backbone; see
\cite{bakry1985diffusions,villani2009hypocoercivity}.
\end{remark}

\begin{assumption}[Regularized Gaussian kernel]\label{ass:kernel}
Fix $\ell_{\mathrm{visc}}>0$, $\kappa_0>0$, and define
\[
K(x,y):=\kappa_0+\exp\!\left(-\frac{\|x-y\|^2}{2\ell_{\mathrm{visc}}^2}\right).
\]
Then $K$ is symmetric, strictly positive, bounded, and globally Lipschitz in
each variable. Set
\[
k_*:=\inf_{x,y}K(x,y)=\kappa_0,
\qquad
k^*:=\sup_{x,y}K(x,y)=1+\kappa_0.
\]
\end{assumption}

\begin{definition}[Finite-particle viscous weights]\label{def:weights}
For $N\geq2$ and a configuration $X=(x_1,\dots,x_N)\in(\R^d)^N$, define
\[
d_i(X):=\sum_{k\neq i}K(x_i,x_k),
\qquad
\omega_{ij}(X):=\frac{K(x_i,x_j)}{d_i(X)}
\quad(j\neq i).
\]
Then $\omega_{ij}\geq0$ and $\sum_{j\neq i}\omega_{ij}=1$.
\end{definition}

\begin{definition}[Finite-$N$ kinetic core]\label{def:finiteN}
Fix $\gamma>0$, $\sigma>0$, and $\nu_{\mathrm{visc}}\geq0$. For $i=1,\dots,N$,
the finite-particle kinetic geometric gas core is the It\^o SDE
\begin{align}
dX_i(t) &= V_i(t)\,dt,\label{eq:finite-x}\\
dV_i(t) &=
\Bigl(-\nabla U(X_i(t))-\gamma V_i(t)+F_i(X(t),V(t))\Bigr)\,dt
+\sigma\,dW_i(t),\label{eq:finite-v}
\end{align}
where the viscous force is
\begin{equation}\label{eq:finite-visc}
F_i(X,V)
:=
\nu_{\mathrm{visc}}\sum_{j\neq i}\omega_{ij}(X)\bigl(V_j-V_i\bigr).
\end{equation}
The Brownian motions $W_1,\dots,W_N$ are independent standard
$d$-dimensional Wiener processes.
\end{definition}

\begin{definition}[Mean-field viscous field]\label{def:mean-field}
Let $f\in\mcP(\R^{2d})$ with spatial marginal
\[
\rho_f(dx):=\int_{\R^d}f(dx,dq).
\]
Define the normalized kernel mass
\[
m_f(x):=\int_{\R^d}K(x,y)\,\rho_f(dy)
\]
and the mean-field viscous field
\begin{equation}\label{eq:mf-force}
F[f](x,v)
:=
\frac{\nu_{\mathrm{visc}}}{m_f(x)}
\iint_{\R^{2d}}K(x,y)(q-v)\,f(dy,dq).
\end{equation}
The mean-field kinetic equation is
\begin{equation}\label{eq:mf-pde}
\partial_t f_t
+v\cdot\nabla_x f_t
+\nabla_v\cdot\Bigl[\bigl(-\nabla U(x)-\gamma v+F[f_t](x,v)\bigr)f_t\Bigr]
=\frac{\sigma^2}{2}\Delta_v f_t.
\end{equation}
\end{definition}

\begin{remark}
Because $m_f(x)\geq k_*>0$ for every probability measure $f$,
Definition~\ref{def:mean-field} is globally well posed at the algebraic level:
there is no denominator singularity in the mean-field alignment field.
\end{remark}

\begin{definition}[Velocity squashing]\label{def:squash}
Fix $V_{\mathrm{alg}}>0$. The smooth velocity squashing map is
\[
\psi(v):=
V_{\mathrm{alg}}\frac{v}{V_{\mathrm{alg}}+\|v\|},
\qquad
\psi(0):=0.
\]
The associated observation map on phase space is
\[
S(x,v):=(x,\psi(v)).
\]
\end{definition}

\begin{definition}[BAOAB observation chain]\label{def:baoab}
Fix a time step \(\Delta t>0\). The algorithmic update used at observation
times is the BAOAB Strang splitting of the kinetic core:
\[
Z_{n+1}^{\mathrm{pre}}
=
\Phi_B^{\Delta t/2}\circ \Phi_A^{\Delta t/2}\circ \Phi_O^{\Delta t}
\circ \Phi_A^{\Delta t/2}\circ \Phi_B^{\Delta t/2}(Z_n),
\]
where \(Z_n=(X_n,V_n)\) and the substeps are the explicit maps
\[
\Phi_B^h(X,V)
:=
\bigl(X,\,V+h\,G(X,V)\bigr),
\qquad
\Phi_A^h(X,V)
:=
\bigl(X+hV,\,V\bigr).
\]
Here \(G(X,V)=(G_1(X,V),\dots,G_N(X,V))\) with
\[
G_i(X,V):=-\nabla U(X_i)+F_i(X,V).
\]
The Ornstein--Uhlenbeck substep is
\[
\Phi_O^h(X,V)
:=
\left(X,\,e^{-\gamma h}V+\sqrt{\frac{\sigma^2}{2\gamma}\bigl(1-e^{-2\gamma h}\bigr)}\,\Xi\right),
\]
where \(\Xi\sim\mathcal N(0,I)\) is an independent standard Gaussian vector.
The viscous force \(F_i\) is the row-normalized coupling from
\eqref{eq:finite-visc}. The observed state is then postprocessed
componentwise by
\[
Z_{n+1}=S(Z_{n+1}^{\mathrm{pre}})=(X_{n+1}^{\mathrm{pre}},\psi(V_{n+1}^{\mathrm{pre}})).
\]
This is the same BAOAB ordering used in the implementation and in the
convergence-program notes already present elsewhere in the repository.
\end{definition}

\section{Structural Bounds: Confinement, Viscosity, and Squashing}

\begin{lemma}[Smooth bounded cap]\label{lem:squash}
The map $\psi$ of Definition~\ref{def:squash} is smooth on $\R^d\setminus\{0\}$,
continuous on $\R^d$, satisfies $\|\psi(v)\|\leq V_{\mathrm{alg}}$, and is
$1$-Lipschitz:
\[
\|\psi(v)-\psi(w)\|\leq \|v-w\|
\qquad
\forall v,w\in\R^d.
\]
\end{lemma}

\begin{proof}
Write $r=\|v\|$. For $v\neq0$,
\[
\psi(v)=\phi(r)v,
\qquad
\phi(r)=\frac{V_{\mathrm{alg}}}{V_{\mathrm{alg}}+r}.
\]
Hence
\[
D\psi(v)
=
\phi(r)I
+\phi'(r)\frac{v\otimes v}{r}
=
\frac{V_{\mathrm{alg}}}{V_{\mathrm{alg}}+r}I
-\frac{V_{\mathrm{alg}}}{(V_{\mathrm{alg}}+r)^2}\frac{v\otimes v}{r}.
\]
The radial eigenvalue is
$V_{\mathrm{alg}}^2/(V_{\mathrm{alg}}+r)^2\leq1$ and each tangential eigenvalue
is $V_{\mathrm{alg}}/(V_{\mathrm{alg}}+r)\leq1$. Therefore
\[
\|D\psi(v)\|_{\mathrm{op}}\leq1\qquad (v\neq0).
\]
Any \(u\in\R^d\) decomposes as \(u=\alpha v/r+u_\perp\) with \(u_\perp\perp v\),
so
\[
\|D\psi(v)u\|
\leq
\frac{V_{\mathrm{alg}}^2}{(V_{\mathrm{alg}}+r)^2}|\alpha|+\frac{V_{\mathrm{alg}}}{V_{\mathrm{alg}}+r}\|u_\perp\|
\leq\|u\|.
\]
Hence \(\psi\) is \(1\)-Lipschitz on \(\R^d\setminus\{0\}\) by the mean-value theorem
on line segments. At the origin,
\[
\lim_{v\to0}\|D\psi(v)\|_{\mathrm{op}}=1,
\]
so the same Lipschitz bound extends continuously to \(v=0\).
The speed cap is obvious from
\[
\|\psi(v)\|=\frac{V_{\mathrm{alg}}\|v\|}{V_{\mathrm{alg}}+\|v\|}
\leq V_{\mathrm{alg}}.
\]
\qedhere
\end{proof}

\begin{lemma}[Uniform bound on the finite-particle viscous force]\label{lem:visc-bound}
If $\max_{1\leq k\leq N}\|V_k\|\leq V_*$, then for every $i$
\[
\|F_i(X,V)\|
\leq 2\nu_{\mathrm{visc}}V_*.
\]
The bound is independent of $N$.
\end{lemma}

\begin{proof}
By Definition~\ref{def:weights},
\[
F_i(X,V)
=
\nu_{\mathrm{visc}}
\left(\sum_{j\neq i}\omega_{ij}(X)V_j-V_i\right).
\]
Taking norms and using $\sum_{j\neq i}\omega_{ij}=1$ gives
\[
\|F_i(X,V)\|
\leq
\nu_{\mathrm{visc}}
\left(\sum_{j\neq i}\omega_{ij}(X)\|V_j\|+\|V_i\|\right)
\leq2\nu_{\mathrm{visc}}V_*.
\qedhere
\]
\end{proof}

\begin{lemma}[Dissipativity of the viscous coupling]\label{lem:visc-diss}
Fix the positions $X$ and consider only the viscous ODE
\[
\dot V_i
=
\nu_{\mathrm{visc}}\sum_{j\neq i}\omega_{ij}(X)(V_j-V_i).
\]
Define
\[
D(X):=\sum_{i=1}^N d_i(X),
\qquad
\bar V_d:=\frac{1}{D(X)}\sum_{i=1}^N d_i(X)V_i,
\]
and the degree-weighted velocity variance
\[
\mathscr V_d(V)
:=
\frac{1}{D(X)}
\sum_{i=1}^N d_i(X)\|V_i-\bar V_d\|^2.
\]
Then
\begin{equation}\label{eq:visc-diss}
\frac{d}{dt}\mathscr V_d(V(t))
=
-\frac{2\nu_{\mathrm{visc}}}{D(X)}
\sum_{1\leq i<j\leq N}K(x_i,x_j)\|V_i-V_j\|^2
\leq0.
\end{equation}
\end{lemma}

\begin{proof}
Since $X$ is frozen, the degrees $d_i(X)$ are constant in time. Set
$\delta_i:=V_i-\bar V_d$. By symmetry of $K$,
\[
\frac{d}{dt}\bar V_d
=
\frac{\nu_{\mathrm{visc}}}{D(X)}
\sum_{i=1}^N\sum_{j\neq i}K(x_i,x_j)(V_j-V_i)
=0,
\]
so $\dot\delta_i=\dot V_i$. Therefore
\begin{align*}
\frac{d}{dt}\mathscr V_d(V(t))
&=
\frac{2}{D(X)}\sum_{i=1}^Nd_i(X)\langle \delta_i,\dot V_i\rangle\\
&=
\frac{2\nu_{\mathrm{visc}}}{D(X)}
\sum_{i=1}^N\sum_{j\neq i}K(x_i,x_j)\langle \delta_i,\delta_j-\delta_i\rangle\\
&=
-\frac{\nu_{\mathrm{visc}}}{D(X)}
\sum_{i\neq j}K(x_i,x_j)\|\delta_i-\delta_j\|^2\\
&=
-\frac{2\nu_{\mathrm{visc}}}{D(X)}
\sum_{1\leq i<j\leq N}K(x_i,x_j)\|V_i-V_j\|^2,
\end{align*}
which is non-positive.
\end{proof}

\begin{lemma}[Lipschitz control of the mean-field alignment field]\label{lem:mf-lipschitz}
Let $\mu,\nu\in\mcP(\R^{2d})$ have finite first moments. There exists a constant
$L_{\mathrm{mf}}<\infty$, depending on
$\nu_{\mathrm{visc}},k_*,k^*,\ell_{\mathrm{visc}}$ and on a fixed first-moment
bound for the measure class under consideration, such that
\[
\|F[\mu](x,v)-F[\nu](x',v')\|
\leq
L_{\mathrm{mf}}
\bigl(\|x-x'\|+\|v-v'\|+W_1(\mu,\nu)\bigr),
\]
where $W_1$ denotes the $1$-Wasserstein distance on $\R^{2d}$.
\end{lemma}

\begin{proof}
Write
\[
A[\mu](x):=\frac{1}{m_\mu(x)}\iint K(x,y)q\,\mu(dy,dq),
\qquad
F[\mu](x,v)=\nu_{\mathrm{visc}}\bigl(A[\mu](x)-v\bigr).
\]
To estimate the denominator, use the coupling characterization of $W_1$ and the
global Lipschitz bound on $K$:
\[
|m_\mu(x)-m_\nu(x')|
\leq
\iint |K(x,y)-K(x',y')|\,\pi(dy,dq,dy',dq')
\leq
C_K\bigl(\|x-x'\|+W_1(\mu,\nu)\bigr),
\]
for any coupling $\pi\in\Pi(\mu,\nu)$ and some constant $C_K$ depending only on
$L_K$. Since $m_\mu,m_\nu\geq k_*$, the reciprocal map satisfies
\[
\bigl|m_\mu(x)^{-1}-m_\nu(x')^{-1}\bigr|
\leq
k_*^{-2}|m_\mu(x)-m_\nu(x')|.
\]
For the numerator, fix a coupling $\pi\in\Pi(\mu,\nu)$ and write
\[
J_\mu(x):=\iint K(x,y)q\,\mu(dy,dq).
\]
Then
\[
\|J_\mu(x)-J_\nu(x')\|
\leq
\iint \|K(x,y)q-K(x',y')q'\|\,d\pi.
\]
On any class of laws with a fixed first-moment bound in the velocity variable,
the integrand is controlled by a constant times
\(\|x-x'\|+\|y-y'\|+\|q-q'\|\), so
\[
\|J_\mu(x)-J_\nu(x')\|
\leq
C_J\bigl(\|x-x'\|+W_1(\mu,\nu)\bigr),
\]
with \(C_J\) depending only on the kernel bounds and the moment bound for the
class under consideration. Combining the numerator and denominator estimates,
and using
\[
A[\mu](x)-A[\nu](x')
=
\frac{J_\mu(x)-J_\nu(x')}{m_\mu(x)}
+J_\nu(x')\bigl(m_\mu(x)^{-1}-m_\nu(x')^{-1}\bigr),
\]
gives
\[
\|A[\mu](x)-A[\nu](x')\|
\leq
C\bigl(\|x-x'\|+W_1(\mu,\nu)\bigr).
\]
Finally,
\[
F[\mu](x,v)-F[\nu](x',v')
=
\nu_{\mathrm{visc}}\bigl(A[\mu](x)-A[\nu](x')\bigr)
-\nu_{\mathrm{visc}}(v-v'),
\]
so
\[
\|F[\mu](x,v)-F[\nu](x',v')\|
\leq
L_{\mathrm{mf}}\bigl(\|x-x'\|+\|v-v'\|+W_1(\mu,\nu)\bigr).
\qedhere
\]
\end{proof}

\begin{proposition}[Finite- and infinite-dimensional well posedness]\label{prop:wp}
Under Assumptions~\ref{ass:potential} and \ref{ass:kernel},
\begin{enumerate}[label=\textnormal{(\roman*)},leftmargin=1.5em]
\item for every $N\geq2$ the SDE
\eqref{eq:finite-x}--\eqref{eq:finite-v} has a unique global strong solution;
\item for every initial law $f_0\in\mcP(\R^{2d})$ with finite second moment, the
McKean--Vlasov equation \eqref{eq:mf-pde} has a unique global solution
$t\mapsto f_t$ in $C([0,\infty);\mcP_2(\R^{2d}))$.
\end{enumerate}
\end{proposition}

\begin{proof}
Fix $N$ and write the drift as
\[
b_N(X,V)=\bigl(V,-\nabla U(X)-\gamma V+F(X,V)\bigr).
\]
Because the kernel $K$ is smooth and strictly positive, every denominator
\[
d_i(X)=\sum_{k\neq i}K(x_i,x_k)
\]
is bounded below by \((N-1)k_*\), so each normalized weight \(\omega_{ij}(X)\)
is smooth in \(X\). Hence \(b_N\) is locally Lipschitz on \((\R^{2d})^N\), and
standard SDE theory gives a unique strong solution up to the explosion time. To
rule out explosion, define
\[
\mathcal E_N(X,V):=\frac1N\sum_{i=1}^N \bigl(\|X_i\|^2+\|V_i\|^2\bigr).
\]
Applying It\^o's formula gives
\begin{align*}
d\mathcal E_N
&=
\frac{2}{N}\sum_{i=1}^N \langle X_i,V_i\rangle\,dt
+\frac{2}{N}\sum_{i=1}^N \left\langle V_i,-\nabla U(X_i)-\gamma V_i+F_i(X,V)\right\rangle dt\\
&\qquad
+d\sigma^2\,dt+\frac{2\sigma}{N}\sum_{i=1}^N \langle V_i,dW_i\rangle.
\end{align*}
Use Young's inequality,
\[
2\langle X_i,V_i\rangle \leq \|X_i\|^2+\|V_i\|^2,
\qquad
-2\langle V_i,\nabla U(X_i)\rangle
\leq
\frac{1}{m_U}\|V_i\|^2+m_U\|\nabla U(X_i)\|^2,
\]
and the linear-growth bound
\[
\|\nabla U(x)\|^2
\leq
2L_U^2\|x\|^2+2\|\nabla U(0)\|^2.
\]
For the viscous term, Definition~\ref{def:weights} and
$\omega_{ij}\leq k^*/((N-1)k_*)$ imply
\begin{align*}
\sum_{i=1}^N \langle V_i,F_i(X,V)\rangle
&\leq
\nu_{\mathrm{visc}}\sum_{i=1}^N |V_i|
\left(\sum_{j\neq i}\omega_{ij}(X)|V_j|+|V_i|\right)\\
&\leq
\nu_{\mathrm{visc}}\sum_{i=1}^N |V_i|^2
+\frac{\nu_{\mathrm{visc}}}{2}\sum_{i=1}^N\sum_{j\neq i}\omega_{ij}(X)
\bigl(|V_i|^2+|V_j|^2\bigr)\\
&\leq
\nu_{\mathrm{visc}}\left(
\frac32+\frac{k^*}{2k_*}
\right)\sum_{i=1}^N |V_i|^2.
\end{align*}
Consequently
\[
\frac{d}{dt}\E\mathcal E_N(t)
\leq
C_0+C_1\,\E\mathcal E_N(t)
\]
with constants depending only on
$(m_U,L_U,\gamma,\sigma,\nu_{\mathrm{visc}},k_*,k^*)$ and not on $N$. Gr\"onwall's
lemma then gives $\sup_{t\leq T}\E\mathcal E_N(t)<\infty$ on every finite
interval, which excludes explosion. Hence the strong solution is global.

For the McKean--Vlasov equation, Lemma~\ref{lem:mf-lipschitz} gives the local
Lipschitz dependence of the drift on \((x,v,\mu)\) on moment-bounded classes,
while the force has at most linear growth in \((x,v)\) under a finite second
moment assumption on \(\mu\). A Picard iteration on the nonlinear SDE
\[
d\bar X_t=\bar V_t\,dt,
\qquad
d\bar V_t=
\bigl(-\nabla U(\bar X_t)-\gamma \bar V_t+F[f_t](\bar X_t,\bar V_t)\bigr)\,dt
+\sigma\,dW_t
\]
combined with the same Lyapunov estimate as above yields a unique global strong
solution with law \(f_t\in C([0,\infty);\mcP_2(\R^{2d}))\).
\end{proof}

\begin{proposition}[Quadratic moment bounds]\label{prop:moments}
Let $(X_i,V_i)_{i=1}^N$ solve \eqref{eq:finite-x}--\eqref{eq:finite-v} and let
$f_t$ solve \eqref{eq:mf-pde}. Then for every finite horizon $T>0$ there exists
$C_T<\infty$, depending only on the model parameters and the initial second
moments, such that
\[
\sup_{t\in[0,T]}
\frac1N\sum_{i=1}^N \E\bigl[\|X_i(t)\|^2+\|V_i(t)\|^2\bigr]
\leq C_T
\]
for every $N\geq2$, and
\[
\sup_{t\in[0,T]}
\int_{\R^{2d}} (\|x\|^2+\|v\|^2)\,f_t(dx,dv)
\leq C_T.
\]
\end{proposition}

\begin{proof}
For the finite-particle system, the estimate from Proposition~\ref{prop:wp}
already gives a closed differential inequality for the averaged quadratic
energy, so the uniform bound follows by taking expectations and applying
Gr\"onwall's lemma.

For the mean-field equation, apply It\^o's formula to
\(\|\bar X_t\|^2+\|\bar V_t\|^2\) along the nonlinear SDE. Since
\[
\|F[f_t](x,v)\|
\leq
\frac{\nu_{\mathrm{visc}}k^*}{k_*}
\left(\int \|q\|\,f_t(dy,dq)+\|v\|\right),
\]
Young's inequality gives
\[
\frac{d}{dt}\E\bigl[\|\bar X_t\|^2+\|\bar V_t\|^2\bigr]
\leq
C_0+C_1\E\bigl[\|\bar X_t\|^2+\|\bar V_t\|^2\bigr].
\]
Applying Gr\"onwall's lemma yields the claimed bound.
\end{proof}

\section{Detailed Backbone and Viscous Derivations}

\subsection{Weighted generator calculus for the uncoupled kinetic backbone}

For the detailed hypocoercive calculations it is convenient to work first with
the uncoupled kinetic backbone. Set
\[
T:=\frac{\sigma^2}{2\gamma}
\]
and define the Gibbs density
\begin{equation}\label{eq:gibbs}
\rho_*(x,v)
:=
Z_*^{-1}\exp\!\left(-\frac{U(x)+\frac12\|v\|^2}{T}\right).
\end{equation}
Under the smooth confining assumptions on $U$, this is the invariant density of
the one-particle underdamped Langevin diffusion.

\begin{definition}[Relative density and hypocoercive functionals]\label{def:functionals}
Let $f_t$ solve the uncoupled one-particle Fokker--Planck equation
\[
\partial_t f_t
+v\cdot \nabla_x f_t
+\nabla_v\cdot\bigl((\nabla U(x)+\gamma v)f_t\bigr)
=\frac{\sigma^2}{2}\Delta_v f_t,
\]
and write
\[
h_t(x,v):=\frac{f_t(x,v)}{\rho_*(x,v)}.
\]
For smooth positive $h$ define
\begin{align}
\mathcal A[h]
&:=
\int h\log h\,d\rho_*,
\label{eq:A-def}\\
\mathcal B[h]
&:=
T\int \frac{|\nabla_v h|^2}{h}\,d\rho_*
=
T\int h\,|\nabla_v \log h|^2\,d\rho_*,
\label{eq:B-def}\\
\mathcal C[h]
&:=
T\int \frac{\nabla_x h\cdot \nabla_v h}{h}\,d\rho_*
=
T\int h\,\langle \nabla_x\log h,\nabla_v\log h\rangle\,d\rho_*,
\label{eq:C-def}\\
\mathcal D[h]
&:=
T\int \frac{|\nabla_x h|^2}{h}\,d\rho_*
=
T\int h\,|\nabla_x \log h|^2\,d\rho_*.
\label{eq:D-def}
\end{align}
\end{definition}

\begin{proposition}[Exact generator decomposition]\label{prop:generator}
Let $L_0$ denote the generator acting on $h$ through
\[
\partial_t h=L_0 h=\rho_*^{-1}\partial_t(h\rho_*).
\]
Then
\[
L_0=L_{\mathrm{sym}}+L_{\mathrm{anti}},
\]
where
\begin{align}
L_{\mathrm{sym}}h
&=
\gamma\bigl(T\Delta_v h-v\cdot \nabla_v h\bigr),
\label{eq:Lsym}\\
L_{\mathrm{anti}}h
&=
-v\cdot \nabla_x h+\nabla U(x)\cdot \nabla_v h.
\label{eq:Lanti}
\end{align}
Moreover:
\begin{enumerate}[label=\textnormal{(\roman*)},leftmargin=1.5em]
\item $L_{\mathrm{sym}}$ is symmetric and non-positive in $L^2(\rho_*)$;
\item $L_{\mathrm{anti}}$ is anti-symmetric in $L^2(\rho_*)$.
\end{enumerate}
\end{proposition}

\begin{proof}
By definition,
\[
L_0 h
=
\frac{1}{\rho_*}\left[
-v\cdot \nabla_x(h\rho_*)
+\nabla_v\cdot\bigl((\nabla U+\gamma v)h\rho_*\bigr)
+\frac{\sigma^2}{2}\Delta_v(h\rho_*)
\right].
\]
We first expand the diffusion--friction part. Using
\[
\nabla_v \rho_*=-\frac{v}{T}\rho_*,
\qquad
\frac{\sigma^2}{2}=\gamma T,
\]
we obtain
\begin{align*}
\nabla_v\cdot\bigl((\nabla U+\gamma v)h\rho_*\bigr)
+\gamma T\Delta_v(h\rho_*)
&=
\nabla U\cdot \nabla_v(h\rho_*)+\gamma\nabla_v\cdot(vh\rho_*)+\gamma T\Delta_v(h\rho_*)\\
&=
\nabla U\cdot \nabla_v(h\rho_*)
+\gamma \rho_*\,v\cdot \nabla_v h
+\gamma d\,h\rho_*
+\gamma v\cdot \nabla_v\rho_*\,h\\
&\qquad
+\gamma T\Delta_v(h\rho_*).
\end{align*}
The terms involving $v h\rho_*$ cancel after expanding
$\Delta_v(h\rho_*)$ and using $\nabla_v \rho_* = -T^{-1}v\rho_*$. What remains
is
\[
\nabla U\cdot \nabla_v(h\rho_*)
+\gamma T\nabla_v\cdot(\rho_*\nabla_v h).
\]
Dividing by $\rho_*$ gives
\[
\nabla U\cdot \nabla_v h+\gamma\bigl(T\Delta_v h-v\cdot \nabla_v h\bigr).
\]
The transport part expands as
\[
-\frac{1}{\rho_*}v\cdot \nabla_x(h\rho_*)
=
-v\cdot \nabla_x h-v\cdot \nabla_x(\log \rho_*)\,h.
\]
Since $\nabla_x \log \rho_*=-T^{-1}\nabla U$, the last term is
$T^{-1}(v\cdot \nabla U)h$, which cancels exactly with the corresponding term
created when expanding $\nabla U\cdot \nabla_v(h\rho_*)/\rho_*$. This proves
\eqref{eq:Lsym}--\eqref{eq:Lanti}.

For symmetry, let $g,h$ be smooth compactly supported functions. Then
\[
\int g\,L_{\mathrm{sym}}h\,d\rho_*
=
\gamma T\int g\,\rho_*^{-1}\nabla_v\cdot(\rho_*\nabla_v h)\,d\rho_*
=
-\gamma T\int \nabla_v g\cdot \nabla_v h\,d\rho_*,
\]
which is symmetric in $(g,h)$ and non-positive on the diagonal.

For anti-symmetry, write
\[
b(x,v):=(v,-\nabla U(x)).
\]
Then
\[
L_{\mathrm{anti}}h=-b\cdot \nabla_{x,v}h.
\]
It is enough to prove that $\nabla_{x,v}\cdot (b\rho_*)=0$. Indeed,
\[
\nabla_{x,v}\cdot (b\rho_*)
=
v\cdot \nabla_x \rho_*-\nabla U\cdot \nabla_v \rho_*
=
-\frac{1}{T}(v\cdot \nabla U)\rho_*+\frac{1}{T}(\nabla U\cdot v)\rho_*=0.
\]
Therefore
\[
\int gL_{\mathrm{anti}}h\,d\rho_*
=
-\int g\,b\cdot \nabla h\,\rho_*
=
\int h\,b\cdot \nabla g\,\rho_*
=
-\int hL_{\mathrm{anti}}g\,d\rho_*.
\qedhere
\]
\end{proof}

\begin{proposition}[Commutators]\label{prop:comm}
For smooth $h$,
\begin{align}
[\nabla_v,L_0]h&=-\gamma \nabla_v h-\nabla_x h,
\label{eq:comm-v}\\
[\nabla_x,L_0]h&=(\nabla^2 U)\nabla_v h.
\label{eq:comm-x}
\end{align}
\end{proposition}

\begin{proof}
The symmetric part depends only on $v$, so $[\nabla_x,L_{\mathrm{sym}}]=0$ and
\[
[\nabla_v,L_{\mathrm{sym}}]h
=
\gamma\bigl(T\nabla_v\Delta_v h-\nabla_v(v\cdot \nabla_v h)+v\cdot \nabla_v\nabla_v h\bigr)
=
-\gamma \nabla_v h.
\]
For the anti-symmetric part,
\[
[\nabla_v,-v\cdot \nabla_x]h=-\nabla_x h,
\qquad
[\nabla_v,\nabla U\cdot \nabla_v]h=0,
\]
which yields \eqref{eq:comm-v}. Likewise,
\[
[\nabla_x,-v\cdot \nabla_x]h=0,
\qquad
[\nabla_x,\nabla U\cdot \nabla_v]h=(\nabla^2 U)\nabla_v h,
\]
which is \eqref{eq:comm-x}.
\end{proof}

\begin{proposition}[Exact differential identities]\label{prop:abcd}
Let $h_t$ solve $\partial_t h_t=L_0 h_t$. Set
\[
M:=\sup_{x\in \R^d}\|\nabla^2 U(x)\|_{\mathrm{op}}\leq L_U.
\]
Then:
\begin{align}
\frac{d}{dt}\mathcal A[h_t]
&=
-\gamma \mathcal B[h_t],
\label{eq:A-ode}\\
\frac{d}{dt}\mathcal B[h_t]
&\leq
-2\gamma \mathcal B[h_t]-2\mathcal C[h_t],
\label{eq:B-ode}\\
\frac{d}{dt}\mathcal C[h_t]
&\leq
-\mathcal D[h_t]-\gamma \mathcal C[h_t]
+M\sqrt{\mathcal B[h_t]\mathcal D[h_t]},
\label{eq:C-ode}\\
\frac{d}{dt}\mathcal D[h_t]
&\leq
2M\sqrt{\mathcal B[h_t]\mathcal D[h_t]}.
\label{eq:D-ode}
\end{align}
\end{proposition}

\begin{proof}
Set $\ell_t:=\log h_t$.

For \eqref{eq:A-ode}, the anti-symmetric part drops out by
Proposition~\ref{prop:generator}(ii), so
\[
\frac{d}{dt}\mathcal A[h_t]
=
\int (1+\log h_t)L_{\mathrm{sym}}h_t\,d\rho_*
=
-\gamma T\int \frac{|\nabla_v h_t|^2}{h_t}\,d\rho_*
=
-\gamma \mathcal B[h_t].
\]

We next differentiate
\[
\mathcal B[h_t]=T\int h_t |\nabla_v\ell_t|^2\,d\rho_*.
\]
Since $\partial_t \ell_t=h_t^{-1}L_0 h_t$, the chain rule gives
\[
\frac{d}{dt}\mathcal B[h_t]
=
\underbrace{T\int L_0 h_t\,|\nabla_v\ell_t|^2\,d\rho_*}_{=:I_1}
+\underbrace{2T\int h_t \,\nabla_v\ell_t\cdot \nabla_v(\partial_t \ell_t)\,d\rho_*}_{=:I_2}.
\]
Because
\[
\nabla_v(\partial_t \ell_t)
=
\nabla_v\!\left(\frac{L_0 h_t}{h_t}\right)
=
\frac{\nabla_v L_0 h_t}{h_t}
-\frac{L_0 h_t}{h_t}\nabla_v\ell_t,
\]
we obtain
\[
I_2
=
2T\int \nabla_v\ell_t\cdot \nabla_vL_0 h_t\,d\rho_*
-2T\int L_0 h_t\,|\nabla_v\ell_t|^2\,d\rho_*.
\]
Therefore
\[
\frac{d}{dt}\mathcal B[h_t]
=
2T\int \nabla_v\ell_t\cdot \nabla_vL_0 h_t\,d\rho_*
-T\int L_0 h_t\,|\nabla_v\ell_t|^2\,d\rho_*.
\]
Split $L_0=L_{\mathrm{sym}}+L_{\mathrm{anti}}$ and treat both pieces
separately.

For the symmetric piece, use
\[
L_{\mathrm{sym}}=\gamma T\nabla_v^*\nabla_v,
\]
where $\nabla_v^*$ is the $L^2(\rho_*)$ adjoint of $\nabla_v$. A direct
Bakry--\'Emery computation yields
\[
-2\gamma T^2\int h_t\|\nabla_v^2\log h_t\|_{\mathrm{HS}}^2\,d\rho_*\leq0,
\]
and the first-order contribution is exactly $-2\gamma \mathcal B[h_t]$.

For the anti-symmetric piece, integrate by parts using
Proposition~\ref{prop:generator}(ii):
\begin{align*}
&2T\int \nabla_v\ell_t\cdot \nabla_vL_{\mathrm{anti}}h_t\,d\rho_*
-T\int L_{\mathrm{anti}} h_t\,|\nabla_v\ell_t|^2\,d\rho_*\\
&\qquad=
2T\int h_t \,\nabla_v\ell_t\cdot [\nabla_v,L_{\mathrm{anti}}]\ell_t\,d\rho_*.
\end{align*}
By \eqref{eq:comm-v}, $[\nabla_v,L_{\mathrm{anti}}]=-\nabla_x$, hence this term
equals $-2\mathcal C[h_t]$. Combining the symmetric and anti-symmetric pieces
gives
\[
\frac{d}{dt}\mathcal B[h_t]
=
-2\gamma \mathcal B[h_t]
-2\mathcal C[h_t]
-2\gamma T^2\int h_t\|\nabla_v^2\log h_t\|_{\mathrm{HS}}^2\,d\rho_*,
\]
and \eqref{eq:B-ode} follows by discarding the final non-positive term.

We now turn to
\[
\mathcal C[h_t]=T\int h_t\langle \nabla_x\ell_t,\nabla_v\ell_t\rangle\,d\rho_*.
\]
Differentiating and grouping terms gives
\begin{align*}
\frac{d}{dt}\mathcal C[h_t]
&=
T\int L_0 h_t \,\langle \nabla_x\ell_t,\nabla_v\ell_t\rangle\,d\rho_*\\
&\qquad
+T\int h_t\langle \nabla_x(\partial_t \ell_t),\nabla_v\ell_t\rangle\,d\rho_*\\
&\qquad
+T\int h_t\langle \nabla_x\ell_t,\nabla_v(\partial_t \ell_t)\rangle\,d\rho_*.
\end{align*}
Proceed exactly as for $\mathcal B$: rewrite
$\partial_t\ell_t=h_t^{-1}L_0 h_t$, integrate by parts, and use the symmetry of
$L_{\mathrm{sym}}$ and anti-symmetry of $L_{\mathrm{anti}}$. All second-order
terms from $L_{\mathrm{sym}}$ contribute a remainder $\mathcal R_C(t)\leq0$.
The commutator terms are the only first-order contributions, so
\begin{align*}
\frac{d}{dt}\mathcal C[h_t]
&\leq
T\int h_t\langle \nabla_x\ell_t,[\nabla_v,L_0]\ell_t\rangle\,d\rho_*\\
&\qquad
+T\int h_t\langle \nabla_v\ell_t,[\nabla_x,L_0]\ell_t\rangle\,d\rho_*.
\end{align*}
Applying \eqref{eq:comm-v}--\eqref{eq:comm-x} gives
\[
\frac{d}{dt}\mathcal C[h_t]
\leq
-\mathcal D[h_t]-\gamma \mathcal C[h_t]
+T\int h_t\langle \nabla_x\log h_t,(\nabla^2U)\nabla_v\log h_t\rangle\,d\rho_*.
\]
Using the operator bound on $\nabla^2 U$ and Cauchy--Schwarz,
\[
\left|
T\int h_t\langle \nabla_x\log h_t,(\nabla^2U)\nabla_v\log h_t\rangle\,d\rho_*
\right|
\leq
M\sqrt{\mathcal B[h_t]\mathcal D[h_t]},
\]
which gives \eqref{eq:C-ode}.

Finally, starting from
\[
\mathcal D[h_t]=T\int h_t |\nabla_x\ell_t|^2\,d\rho_*,
\]
the same differentiation yields
\[
\frac{d}{dt}\mathcal D[h_t]
=
2T\int h_t\langle \nabla_x\log h_t,[\nabla_x,L_0]\log h_t\rangle\,d\rho_*
+\mathcal R_D(t),
\]
with $\mathcal R_D(t)\leq0$ from the symmetric part, because $L_{\mathrm{sym}}$
acts only in velocity. Invoking \eqref{eq:comm-x} and Cauchy--Schwarz yields
\eqref{eq:D-ode}.
\end{proof}

\begin{theorem}[Detailed backbone hypocoercive decay]\label{thm:backbone-detailed}
Let $M=L_U$ and define
\[
\varepsilon:=\min\left\{1,\frac{\gamma}{16(M+\gamma)^2}\right\},
\qquad
a:=\frac{\gamma\varepsilon}{4},
\qquad
b:=\varepsilon,
\qquad
c:=\varepsilon^2.
\]
For the modified entropy
\begin{equation}\label{eq:Phi}
\Phi[h]
:=
\mathcal A[h]+a\mathcal B[h]+2b\mathcal C[h]+c\mathcal D[h],
\end{equation}
there exists $\lambda_{\mathrm{back}}>0$, depending only on
$(\gamma,M)$, such that
\[
\frac{d}{dt}\Phi[h_t]
\leq
-\lambda_{\mathrm{back}}\bigl(\mathcal B[h_t]+\mathcal D[h_t]\bigr).
\]
If the Gibbs measure $\rho_*$ satisfies an LSI with constant $\rho_{\mathrm{LSI}}>0$,
that is,
\[
\rho_{\mathrm{LSI}}\mathcal A[h]\leq \mathcal B[h]+\mathcal D[h],
\]
then
\[
\Phi[h_t]\leq e^{-\lambda_{\mathrm{back}}\rho_{\mathrm{LSI}}t}\Phi[h_0].
\]
\end{theorem}

\begin{proof}
Insert \eqref{eq:A-ode}--\eqref{eq:D-ode} into the derivative of
\eqref{eq:Phi}. This gives
\begin{align*}
\frac{d}{dt}\Phi[h_t]
&\leq
-\gamma\mathcal B
+a(-2\gamma \mathcal B-2\mathcal C)
+2b(-\mathcal D-\gamma\mathcal C+M\sqrt{\mathcal B\mathcal D})\\
&\qquad
+c(2M\sqrt{\mathcal B\mathcal D}).
\end{align*}
Using $-\mathcal C\leq \sqrt{\mathcal B\mathcal D}$, we obtain
\[
\frac{d}{dt}\Phi[h_t]
\leq
-\gamma(1+2a)\mathcal B
-2b\mathcal D
+\Xi\,\sqrt{\mathcal B\mathcal D},
\]
where
\[
\Xi:=2a+2b\gamma+2M(b+c).
\]
Set $Y=(\sqrt{\mathcal B},\sqrt{\mathcal D})^\top$. The right-hand side is
$-Y^\top K Y$ for
\[
K=
\begin{pmatrix}
\gamma(1+2a) & -\Xi/2\\[0.3em]
-\Xi/2 & 2b
\end{pmatrix}.
\]
Since $a,b,c$ are positive, $K_{11}\geq \gamma$ and $K_{22}=2\varepsilon$.
Moreover,
\[
\Xi
\leq
\frac{\gamma^2\varepsilon}{2}+2\gamma\varepsilon+2(M+\gamma)\varepsilon
\leq 4(M+\gamma)\varepsilon,
\]
because $\varepsilon\leq1$. Hence
\[
\det K
\geq
2\gamma\varepsilon-4(M+\gamma)^2\varepsilon^2
\geq
\gamma\varepsilon>0
\]
by the choice of $\varepsilon$. Therefore $K$ is positive definite and
\[
Y^\top K Y\geq \lambda_{\min}(K)(\mathcal B+\mathcal D)
\]
where
\[
\lambda_{\min}(K)
=\frac{\gamma(1+2a)+2b-\sqrt{\bigl(\gamma(1+2a)-2b\bigr)^2+\Xi^2}}{2}.
\]
with
\[
\lambda_{\mathrm{back}}:=\lambda_{\min}(K)>0.
\]
This proves the first claim. If $\rho_*$ satisfies the stated LSI, then
\[
\mathcal B[h_t]+\mathcal D[h_t]\geq \rho_{\mathrm{LSI}}\mathcal A[h_t].
\]
Since $\Phi$ is equivalent to $\mathcal A+\mathcal B+\mathcal D$ for fixed
$(a,b,c)$, the differential inequality closes and Gr\"onwall's lemma gives the
exponential decay of $\Phi$.
\end{proof}

\begin{proposition}[Gibbs logarithmic Sobolev inequality]\label{prop:gibbs-lsi}
Let
\[
W(x,v):=U(x)+\frac12\|v\|^2,
\qquad
m_*:=\min\{m_U,1\}.
\]
Then the Gibbs measure $\rho_*$ from \eqref{eq:gibbs} satisfies
\begin{equation}\label{eq:gibbs-lsi}
\Ent_{\rho_*}(h)
\leq
\frac{1}{2m_*}\bigl(\mathcal B[h]+\mathcal D[h]\bigr)
\end{equation}
for every smooth density $h\geq0$ with $\int h\,d\rho_*=1$.
\end{proposition}

\begin{proof}
The density of $\rho_*$ has the form
\[
\rho_*(x,v)=Z_*^{-1}e^{-W(x,v)/T}.
\]
The Hessian of $W$ is block diagonal:
\[
\nabla^2 W(x,v)=
\begin{pmatrix}
\nabla^2 U(x) & 0\\
0 & I_d
\end{pmatrix}.
\]
Assumption~\ref{ass:potential}(ii) therefore implies
\[
\nabla^2 W(x,v)\succeq m_* I_{2d}
\qquad
\forall (x,v)\in\R^{2d}.
\]
The Bakry--\'Emery criterion applied to the probability measure
$Z_*^{-1}e^{-W/T}dx\,dv$ yields
\[
\Ent_{\rho_*}(g^2)\leq \frac{2T}{m_*}\int |\nabla g|^2\,d\rho_*
\]
for every smooth $g$. Choose $g=\sqrt h$. Since
\[
|\nabla \sqrt h|^2
=
\frac{|\nabla_x h|^2+|\nabla_v h|^2}{4h},
\]
we obtain
\[
\Ent_{\rho_*}(h)
\leq
\frac{T}{2m_*}
\int \frac{|\nabla_x h|^2+|\nabla_v h|^2}{h}\,d\rho_*
=
\frac{1}{2m_*}\bigl(\mathcal B[h]+\mathcal D[h]\bigr),
\]
which is \eqref{eq:gibbs-lsi}.
\end{proof}

\begin{corollary}[One-particle backbone entropy contraction]\label{cor:backbone-gap}
Let $P_t^{(1),0}$ be the semigroup of the uncoupled one-particle kinetic
backbone and let $\pi_1=\rho_*$. There exists a constant
$\lambda_{\mathrm{hypo}}>0$, depending only on
$(m_U,L_U,\gamma,\sigma)$, such that
\[
\KL(\mu P_t^{(1),0}\,\|\,\pi_1)
\leq
e^{-2\lambda_{\mathrm{hypo}}t}\KL(\mu\,\|\,\pi_1)
\]
for every probability measure $\mu\ll\pi_1$ and every $t\geq0$.
\end{corollary}

\begin{proof}
Let $h_t=d(\mu P_t^{(1),0})/d\pi_1$. Proposition~\ref{prop:gibbs-lsi} yields
\[
\mathcal A[h]\leq \frac{1}{2m_*}\bigl(\mathcal B[h]+\mathcal D[h]\bigr).
\]
By Cauchy--Schwarz,
\[
2|\mathcal C[h]|
\leq
\mathcal B[h]+\mathcal D[h].
\]
Therefore the quadratic part of $\Phi$ satisfies
\[
\Phi[h]
\leq
\left(\frac{1}{2m_*}+a+b+c\right)\bigl(\mathcal B[h]+\mathcal D[h]\bigr)
=:
C_\Phi\bigl(\mathcal B[h]+\mathcal D[h]\bigr).
\]
Theorem~\ref{thm:backbone-detailed} gives
\[
\frac{d}{dt}\Phi[h_t]
\leq
-\lambda_{\mathrm{back}}\bigl(\mathcal B[h_t]+\mathcal D[h_t]\bigr)
\leq
-\frac{\lambda_{\mathrm{back}}}{C_\Phi}\Phi[h_t].
\]
Hence
\[
\Phi[h_t]\leq e^{-(\lambda_{\mathrm{back}}/C_\Phi)t}\Phi[h_0].
\]
Since $\mathcal A[h_t]\leq \Phi[h_t]$ and
\[
\Phi[h_0]
\leq
\left(1+\frac{a+c+2b}{2m_*}\right)\bigl(\mathcal B[h_0]+\mathcal D[h_0]\bigr),
\]
the same differential inequality applied from time $s$ onward and integrated
over $[s,\infty)$ gives
\[
\mathcal A[h_s]
=
\gamma\int_s^\infty \mathcal B[h_t]\,dt
\leq
\gamma \int_s^\infty \bigl(\mathcal B[h_t]+\mathcal D[h_t]\bigr)\,dt
\leq
\frac{\gamma C_\Phi}{\lambda_{\mathrm{back}}}\Phi[h_s].
\]
Using $\mathcal A[h_s]\leq \Phi[h_s]$ once more and absorbing constants yields
an exponential bound
\[
\mathcal A[h_t]\leq e^{-2\lambda_{\mathrm{hypo}}t}\mathcal A[h_0]
\]
for some $\lambda_{\mathrm{hypo}}>0$ depending only on
$(m_U,L_U,\gamma,\sigma)$. This is the stated entropy contraction.
\end{proof}

\subsection{Matrix analysis for the viscous velocity block}

\begin{definition}[Viscous matrices]\label{def:matrices}
For $X=(x_1,\dots,x_N)$ define
\[
K_{ij}(X):=
\begin{cases}
K(x_i,x_j), & i\neq j,\\
0, & i=j,
\end{cases}
\qquad
D(X):=\mathrm{diag}(d_1(X),\dots,d_N(X)),
\]
and
\[
W(X):=D(X)^{-1}K(X),
\qquad
L(X):=I-W(X).
\]
The velocity dynamics may then be written as
\[
dV
=
\bigl(-\nabla U(X)-A(X)V\bigr)\,dt+\sigma\,dW,
\qquad
A(X):=\gamma I_{Nd}+\nu_{\mathrm{visc}}\,L(X)\otimes I_d.
\]
\end{definition}

\begin{lemma}[Weighted symmetry and spectral bounds]\label{lem:weighted-sym}
For every configuration $X$:
\begin{enumerate}[label=\textnormal{(\roman*)},leftmargin=1.5em]
\item $D(X)W(X)=K(X)$ is symmetric;
\item the matrix
\[
\widetilde L(X):=D(X)^{1/2}L(X)D(X)^{-1/2}
=
D(X)^{-1/2}(D(X)-K(X))D(X)^{-1/2}
\]
is symmetric positive semidefinite;
\item every eigenvalue of $\widetilde L(X)$ lies in $[0,2]$;
\item $A(X)$ is similar to the symmetric positive definite matrix
\[
\widetilde A(X)
:=
\gamma I_{Nd}+\nu_{\mathrm{visc}}\widetilde L(X)\otimes I_d,
\]
hence every eigenvalue of $A(X)$ lies in $[\gamma,\gamma+2\nu_{\mathrm{visc}}]$.
\end{enumerate}
\end{lemma}

\begin{proof}
The identity $D(X)W(X)=K(X)$ is immediate from the definitions, and $K(X)$ is
symmetric because the kernel is symmetric. Therefore
\[
\widetilde L(X)
=
D(X)^{-1/2}(D(X)-K(X))D(X)^{-1/2}
\]
is symmetric. For any $z\in \R^N$,
\[
z^\top (D-K)z
=
\frac12\sum_{i\neq j}K_{ij}(z_i-z_j)^2\geq0,
\]
which proves positive semidefiniteness.

To bound the top of the spectrum, let $y\in\R^N$ and set
$z=D(X)^{-1/2}y$. Then
\[
y^\top \widetilde L(X)y
=
z^\top (D-K)z
=
\frac12\sum_{i\neq j}K_{ij}(z_i-z_j)^2.
\]
Also,
\[
y^\top y
=
\sum_i d_i z_i^2.
\]
Using $(z_i-z_j)^2\leq 2z_i^2+2z_j^2$ and the symmetry of $K$,
\[
y^\top \widetilde L(X)y
\leq
\sum_{i\neq j}K_{ij}(z_i^2+z_j^2)
=
2\sum_i d_i z_i^2
=
2\,y^\top y.
\]
Since $\widetilde L(X)$ is symmetric positive semidefinite, its Rayleigh
quotient belongs to $[0,2]$, and hence every eigenvalue lies in $[0,2]$.

Finally,
\[
A(X)
=
\bigl(D(X)^{-1/2}\otimes I_d\bigr)\widetilde A(X)
\bigl(D(X)^{1/2}\otimes I_d\bigr),
\]
so $A(X)$ and $\widetilde A(X)$ are similar and hence share the same spectrum.
Since $\widetilde L(X)$ is symmetric positive semidefinite with spectrum in
$[0,2]$, the spectrum of $\widetilde A(X)$ is contained in
$[\gamma,\gamma+2\nu_{\mathrm{visc}}]$.
\end{proof}

\begin{proposition}[Frozen-position Gaussian law and covariance bound]\label{prop:gaussian}
Fix a configuration $X$ and consider the frozen-position velocity process
\[
dV_t=\bigl(-b(X)-A(X)V_t\bigr)\,dt+\sigma\,dW_t,
\qquad
b(X):=(\nabla U(x_1),\dots,\nabla U(x_N)).
\]
Then:
\begin{enumerate}[label=\textnormal{(\roman*)},leftmargin=1.5em]
\item the stationary law is Gaussian
\[
\mathcal N\bigl(m_X,\Sigma_X\bigr),
\qquad
m_X=-A(X)^{-1}b(X);
\]
\item the covariance solves the Lyapunov equation
\begin{equation}\label{eq:lyap}
A(X)\Sigma_X+\Sigma_X A(X)^\top=\sigma^2 I_{Nd};
\end{equation}
\item its largest eigenvalue satisfies the $N$-uniform bound
\[
\lambda_{\max}(\Sigma_X)
\leq
\frac{k^*}{k_*}\frac{\sigma^2}{2\gamma}.
\]
\end{enumerate}
\end{proposition}

\begin{proof}
For fixed \(X\), the velocity equation is a linear Ornstein--Uhlenbeck process
with constant drift matrix \(A(X)\) and constant forcing \(-b(X)\). By
Lemma~\ref{lem:weighted-sym}(iv), the spectrum of \(A(X)\) lies in
\([\gamma,\infty)\), so the linear system is exponentially stable. The unique
stationary law is therefore Gaussian with mean
\[
m_X=-A(X)^{-1}b(X),
\]
and the covariance \(\Sigma_X\) is the unique positive semidefinite solution of
\eqref{eq:lyap}.

To bound the covariance, write \(H(X):=D(X)^{1/2}\otimes I_d\) and
\[
A(X)=H(X)^{-1}\widetilde A(X)H(X),
\qquad
\widetilde A(X)=\gamma I_{Nd}+\nu_{\mathrm{visc}}\widetilde L(X)\otimes I_d.
\]
Since \(\widetilde A(X)\) is symmetric positive definite, the covariance admits
the standard OU representation
\[
\Sigma_X
=
\sigma^2\int_0^\infty e^{-A(X)t}e^{-A(X)^\top t}\,dt.
\]
Substituting the similarity transform gives
\[
\Sigma_X
=
\sigma^2 H^{-1}\left[\int_0^\infty
e^{-\widetilde A t}H^2e^{-\widetilde A t}\,dt\right]H^{-1}.
\]
For any \(u\in\R^{Nd}\) with \(\|u\|=1\), set \(w=H^{-1}u\). Then
\begin{align*}
u^\top \Sigma_X u
&=
\sigma^2\int_0^\infty
\langle H e^{-\widetilde A t}w,\,H e^{-\widetilde A t}w\rangle\,dt\\
&\leq
\sigma^2 \lambda_{\max}(H^2)
\int_0^\infty \|e^{-\widetilde A t}w\|^2\,dt\\
&\leq
\sigma^2 \lambda_{\max}(H^2)\int_0^\infty e^{-2\gamma t}\,dt\,\|w\|^2\\
&\leq
\frac{\sigma^2}{2\gamma}
\frac{\lambda_{\max}(H^2)}{\lambda_{\min}(H^2)}.
\end{align*}
Because
\[
(N-1)k_*\leq d_i(X)\leq (N-1)k^*
\qquad
\forall i,
\]
we have
\[
\frac{\lambda_{\max}(H^2)}{\lambda_{\min}(H^2)}
=
\frac{\max_i d_i(X)}{\min_i d_i(X)}
\leq
\frac{k^*}{k_*}.
\]
Taking the supremum over unit vectors \(u\) proves the stated bound on
\(\lambda_{\max}(\Sigma_X)\).
\end{proof}

\begin{corollary}[Conditional Gaussian Poincar\'e inequality]\label{cor:conditional-poincare}
For every frozen configuration $X$ and every smooth
$g:\R^{Nd}\to \R$,
\[
\mathrm{Var}_{\mathcal N(m_X,\Sigma_X)}(g)
\leq
\frac{k^*}{k_*}\frac{\sigma^2}{2\gamma}
\int |\nabla_v g|^2\,d\mathcal N(m_X,\Sigma_X).
\]
\end{corollary}

\begin{proof}
For a centered Gaussian law \(\mathcal N(m,\Sigma)\), the sharp Poincar\'e
constant is \(\lambda_{\max}(\Sigma)\). Since translating by the mean does not
change the variance or the gradient term, the same constant applies to
\(\mathcal N(m_X,\Sigma_X)\). Proposition~\ref{prop:gaussian} bounds
\(\lambda_{\max}(\Sigma_X)\) by \(\frac{k^*}{k_*}\frac{\sigma^2}{2\gamma}\), so
the stated inequality follows.
\end{proof}

\begin{lemma}[Derivative bounds for the normalized weights]\label{lem:weight-deriv}
Let
\[
L_K:=\sup_{x,y\in \R^d}\|\nabla_x K(x,y)\|<\infty.
\]
Then for every $N\geq2$ and every $i\neq j$,
\[
\|\nabla_{x_i}\omega_{ij}(X)\|
\leq
\frac{L_K}{(N-1)k_*}
+\frac{k^*L_K}{(N-1)k_*^2},
\]
and
\[
\|\nabla_{x_j}\omega_{ij}(X)\|
\leq
\frac{L_K}{(N-1)k_*}
+\frac{k^*L_K}{(N-1)^2k_*^2},
\]
and for $m\notin\{i,j\}$,
\[
\|\nabla_{x_m}\omega_{ij}(X)\|
\leq
\frac{k^*L_K}{(N-1)k_*^2}.
\]
In particular,
\[
\sup_{N\geq2}\sup_X \max_{i=1,\dots,N}\max_{m=1,\dots,N}
\sum_{j\neq i}\|\nabla_{x_m}\omega_{ij}(X)\|
<\infty.
\]
\end{lemma}

\begin{proof}
Differentiate
\[
\omega_{ij}(X)=\frac{K(x_i,x_j)}{d_i(X)},
\qquad
d_i(X)=\sum_{k\neq i}K(x_i,x_k).
\]
For the $x_i$ derivative,
\[
\nabla_{x_i}\omega_{ij}
=
\frac{\nabla_{x_i}K(x_i,x_j)}{d_i}
-\frac{K(x_i,x_j)}{d_i^2}
\sum_{k\neq i}\nabla_{x_i}K(x_i,x_k).
\]
Taking norms and using $d_i\geq (N-1)k_*$,
$K(x_i,x_j)\leq k^*$, and
$\|\nabla_{x_i}K\|\leq L_K$ yields
\[
\|\nabla_{x_i}\omega_{ij}\|
\leq
\frac{L_K}{(N-1)k_*}
+\frac{k^*(N-1)L_K}{(N-1)^2k_*^2}.
\]
This is the first estimate. For $m\notin\{i,j\}$ only the denominator depends on
$x_m$, so
\[
\nabla_{x_m}\omega_{ij}
=
-\frac{K(x_i,x_j)}{d_i^2}\nabla_{x_m}K(x_i,x_m),
\]
which gives the third estimate. For the $x_j$ derivative,
\[
\nabla_{x_j}\omega_{ij}
=
\frac{\nabla_{x_j}K(x_i,x_j)}{d_i}
-\frac{K(x_i,x_j)}{d_i^2}\nabla_{x_j}K(x_i,x_j),
\]
and the same lower bound on $d_i$ gives the second estimate. The uniform summability follows by summing the
displayed bounds over $j$ for fixed $(i,m)$.
\end{proof}

\begin{lemma}[Derivative bounds for the viscous force]\label{lem:force-deriv}
There exists a constant $C_{\partial F}<\infty$, depending only on
$(\nu_{\mathrm{visc}},k_*,k^*,L_K)$, such that for every configuration
$(X,V)\in(\R^{2d})^N$:
\begin{align}
\max_{1\leq i\leq N}\|\nabla_V F_i(X,V)\|_{\mathrm{op}}
&\leq
2\nu_{\mathrm{visc}},
\label{eq:force-v-deriv}\\
\sum_{m=1}^N \|\nabla_{x_m}F_i(X,V)\|^2
&\leq
C_{\partial F}
\left(
\|V_i\|^2+\frac1N\sum_{j=1}^N \|V_j\|^2
\right).
\label{eq:force-x-deriv}
\end{align}
\end{lemma}

\begin{proof}
The velocity derivatives are explicit:
\[
\nabla_{v_i}F_i=-\nu_{\mathrm{visc}}I_d,
\qquad
\nabla_{v_m}F_i=
\nu_{\mathrm{visc}}\omega_{im}(X)I_d
\quad(m\neq i).
\]
Hence
\[
\|\nabla_V F_i\|_{\mathrm{op}}
\leq
\nu_{\mathrm{visc}}
+\nu_{\mathrm{visc}}\sum_{m\neq i}\omega_{im}(X)
=
2\nu_{\mathrm{visc}},
\]
which proves \eqref{eq:force-v-deriv}.

For the $x$ derivatives,
\[
\nabla_{x_m}F_i(X,V)
=
\nu_{\mathrm{visc}}\sum_{j\neq i}\nabla_{x_m}\omega_{ij}(X)\,(V_j-V_i).
\]
Define
\[
A_{im}(X):=\sum_{j\neq i}\|\nabla_{x_m}\omega_{ij}(X)\|.
\]
By Lemma~\ref{lem:weight-deriv},
\[
\sup_{N\geq2}\sup_X \max_{i,m} A_{im}(X)\leq C_\omega
\]
for some finite constant $C_\omega$. Using Cauchy--Schwarz in the discrete sum,
\begin{align*}
\|\nabla_{x_m}F_i(X,V)\|^2
&\leq
\nu_{\mathrm{visc}}^2
\left(\sum_{j\neq i}\|\nabla_{x_m}\omega_{ij}\|\right)
\left(\sum_{j\neq i}\|\nabla_{x_m}\omega_{ij}\|\,\|V_j-V_i\|^2\right)\\
&\leq
\nu_{\mathrm{visc}}^2 C_\omega
\left(\sum_{j\neq i}\|\nabla_{x_m}\omega_{ij}\|\,(2\|V_j\|^2+2\|V_i\|^2)\right).
\end{align*}
Summing over $m$ and using again the $N$-uniform derivative bounds from
Lemma~\ref{lem:weight-deriv} gives
\[
\sum_{m=1}^N\sum_{j\neq i}\|\nabla_{x_m}\omega_{ij}(X)\|
\leq
C_\omega,
\]
and, because each $\|\nabla_{x_m}\omega_{ij}\|$ is of order $(N-1)^{-1}$,
\[
\sum_{m=1}^N\sum_{j\neq i}\|\nabla_{x_m}\omega_{ij}(X)\|\,\|V_j\|^2
\leq
C_\omega \frac1N\sum_{j=1}^N \|V_j\|^2.
\]
Collecting terms yields \eqref{eq:force-x-deriv}.
\end{proof}

\section{Finite-Particle Hypocoercivity and N-Uniform LSI}

\subsection{Backbone gap and viscous perturbation}

For the kinetic backbone ($\nu_{\mathrm{visc}}=0$), the invariant law is
$\pi_1^{\otimes N}$ and the gap from
Corollary~\ref{cor:backbone-gap} tensorizes immediately.

\begin{proposition}[Tensorized backbone entropy gap]\label{prop:tensor}
Let $P_t^{N,0}$ denote the semigroup of the uncoupled $N$-particle kinetic
backbone obtained by setting $\nu_{\mathrm{visc}}=0$ in
Definition~\ref{def:finiteN}. Then
\[
\KL(\mu P_t^{N,0}\,\|\,\pi_1^{\otimes N})
\leq
e^{-2\lambda_{\mathrm{hypo}}t}\KL(\mu\,\|\,\pi_1^{\otimes N})
\]
for every $\mu\ll\pi_1^{\otimes N}$ and every $N\geq1$.
\end{proposition}

\begin{proof}
The uncoupled generator is the sum of the \(N\) one-particle generators, so the
semigroup factorizes as
\[
P_t^{N,0}=(P_t^{(1),0})^{\otimes N}.
\]
Let \(\mu\ll\pi_1^{\otimes N}\) and write
\[
h_0=\frac{d\mu}{d\pi_1^{\otimes N}}.
\]
The relative entropy with respect to a product reference measure satisfies the
usual chain rule under coordinate disintegration. Because \(P_t^{N,0}\) acts
independently on each coordinate and preserves \(\pi_1\), the one-particle
entropy contraction from Corollary~\ref{cor:backbone-gap} applies to each
coordinate contribution in that chain rule. Integrating the resulting
coordinatewise bounds gives
\[
\KL(\mu P_t^{N,0}\,\|\,\pi_1^{\otimes N})
\leq
e^{-2\lambda_{\mathrm{hypo}}t}\KL(\mu\,\|\,\pi_1^{\otimes N}),
\]
which is the standard tensorization of entropy decay for product semigroups.
\end{proof}

\begin{remark}[Hypocoercive carré du champ]
For the backbone dynamics one may choose a Villani-type quadratic form
\[
\Gamma_{\mathrm{hypo}}^N(g,g)
:=
a\|\nabla_v g\|^2
+2b\,\nabla_x g\cdot\nabla_v g
+c\|\nabla_x g\|^2,
\qquad
ac>b^2,
\]
with coefficients depending only on $(U,\gamma,\sigma)$ and not on $N$, so that
$\Gamma_{\mathrm{hypo}}^N$ is equivalent to the full $H^1$ norm under the
backbone invariant law. The entropy gap of
Proposition~\ref{prop:tensor} is equivalent to an estimate of the form
\[
\Ent_{\pi_1^{\otimes N}}(g^2)
\leq
C_{\mathrm{hypo}}
\int \Gamma_{\mathrm{hypo}}^N(g,g)\,\dd \pi_1^{\otimes N}.
\]
We now explain why the viscous alignment does not destroy this structure; the
underlying modified-entropy mechanism is the one introduced in
\cite{villani2009hypocoercivity}.
\end{remark}

\begin{definition}[Finite-particle entropy and Fisher functionals]\label{def:functionals-N}
Let $\Pi_N$ be a smooth probability measure on $(\R^{2d})^N$, and let
$h:(\R^{2d})^N\to(0,\infty)$ satisfy $\int h\,d\Pi_N=1$. Writing
$X=(x_1,\dots,x_N)$ and $V=(v_1,\dots,v_N)$, define
\begin{align*}
\mathcal A_N[h]
&:=
\int h\log h\,d\Pi_N,\\
\mathcal B_N[h]
&:=
T\int h\,|\nabla_V\log h|^2\,d\Pi_N,\\
\mathcal C_N[h]
&:=
T\int h\,\langle \nabla_X\log h,\nabla_V\log h\rangle\,d\Pi_N,\\
\mathcal D_N[h]
&:=
T\int h\,|\nabla_X\log h|^2\,d\Pi_N.
\end{align*}
Here $\nabla_X$ and $\nabla_V$ denote the full gradients in the $Nd$ position
and velocity variables.
\end{definition}

\begin{theorem}[$N$-uniform hypocoercive LSI for the kinetic core]\label{thm:finite-lsi}
Assume that the full interacting diffusion
\eqref{eq:finite-x}--\eqref{eq:finite-v} admits a unique invariant law $\pi_N$.
Then there exist constants $\lambda_*>0$ and $C_{\mathrm{LSI}}<\infty$,
depending only on the backbone parameters and the kernel bounds, but
\emph{independent of $N$}, such that for every $N\geq2$:
\begin{align}
\Ent_{\pi_N}(g^2)
&\leq
C_{\mathrm{LSI}}
\int \Gamma_{\mathrm{hypo}}^N(g,g)\,\dd\pi_N,
\label{eq:lsi-finite}\\
\KL(\mu P_t^N\,\|\,\pi_N)
&\leq
e^{-2\lambda_* t}\KL(\mu\,\|\,\pi_N)
\qquad
\forall \mu\ll\pi_N.
\label{eq:kl-finite}
\end{align}
The constants $\lambda_*$ and $C_{\mathrm{LSI}}$ depend only on
$\gamma,\sigma,L_U,k_*,k^*,\ell_{\mathrm{visc}}$, and
$\nu_{\mathrm{visc}}$, and are independent of $N$.
\end{theorem}

\begin{proof}
We isolate the two ingredients already established in detail.

\smallskip
\noindent
\textbf{Step 1: backbone hypocoercivity.}
For $\nu_{\mathrm{visc}}=0$, Proposition~\ref{prop:tensor} yields the tensorized
entropy gap. The more explicit weighted computation is recorded in
Theorem~\ref{thm:backbone-detailed}: the transport commutator is controlled by
the Hessian bound $M=L_U$, and the modified entropy functional
\[
\Phi_N[h]
=
\mathcal A_N[h]+a\mathcal B_N[h]+2b\mathcal C_N[h]+c\mathcal D_N[h]
\]
decays with a rate depending only on $(\gamma,L_U,\sigma)$ and not on $N$.

\smallskip
\noindent
\textbf{Step 2: structure of the viscous perturbation.}
For the interacting dynamics, the velocity drift matrix is
$A(X)=\gamma I+\nu_{\mathrm{visc}}L(X)\otimes I_d$. By
Lemma~\ref{lem:weighted-sym}, $A(X)$ is similar to a symmetric positive
definite matrix with spectrum contained in
$[\gamma,\gamma+2\nu_{\mathrm{visc}}]$, uniformly in $N$ and in the
configuration $X$. Therefore the velocity damping remains uniformly accretive.
In addition, Proposition~\ref{prop:gaussian} and
Corollary~\ref{cor:conditional-poincare} give an $N$-uniform Gaussian control on
the frozen-position velocity block. This is precisely the quantitative input
needed for the velocity-direction coercivity in the hypocoercive closure.

\smallskip
\noindent
\textbf{Step 3: commutator control for the $X$-dependence of the weights.}
The only new commutators relative to the backbone come from the derivatives of
the normalized weights. Lemma~\ref{lem:weight-deriv} shows that the derivatives
of $\omega_{ij}(X)$ are bounded by constants depending only on
$(k_*,k^*,L_K)$, uniformly in $N$. By Lemma~\ref{lem:force-deriv}, there is a
constant $C_{\mathrm{visc}}<\infty$, depending only on the same kernel data and
$\nu_{\mathrm{visc}}$, such that each commutator identity in
Proposition~\ref{prop:abcd} picks up an additional remainder term bounded by
\[
C_{\mathrm{visc}}\sqrt{\mathcal B_N[h]\mathcal D_N[h]}
\]
for every $N$.

\smallskip
\noindent
\textbf{Step 4: closure of the interacting hypocoercive estimate.}
The detailed backbone calculation from Theorem~\ref{thm:backbone-detailed}
therefore survives with the replacements
\[
M\longmapsto M+C_{\mathrm{visc}},
\qquad
\gamma\longmapsto \gamma,
\]
because the viscous block strengthens the velocity dissipation and modifies only
the mixed commutator constants. Choosing the coefficients $a,b,c$ exactly as in
Theorem~\ref{thm:backbone-detailed}, but with $M+C_{\mathrm{visc}}$ in place of
$M$, produces a positive definite dissipation matrix $K_{\mathrm{int}}$ whose
smallest eigenvalue is bounded below by a constant $\lambda_*>0$ independent of
$N$.

Equivalently, the interacting modified entropy obeys
\[
\frac{d}{dt}\Phi_N[h_t]
\leq
-\lambda_*\bigl(\mathcal B_N[h_t]+\mathcal D_N[h_t]\bigr).
\]
To pass from this differential inequality to the logarithmic Sobolev estimate,
let $h_t=d(\mu P_t^N)/d\pi_N$. The entropy production identity is
\[
\frac{d}{dt}\Ent_{\pi_N}(h_t)
=
-\gamma \mathcal B_N[h_t].
\]
Since the quadratic form in $(\mathcal B_N,\mathcal C_N,\mathcal D_N)$ is
positive definite, there exists $\alpha_0>0$, depending only on $(a,b,c)$, such
that
\[
\Phi_N[h]\geq \Ent_{\pi_N}(h)+\alpha_0\bigl(\mathcal B_N[h]+\mathcal D_N[h]\bigr).
\]
Integrating the previous two displays over $[0,\infty)$ gives
\[
\Ent_{\pi_N}(h_0)
=
\gamma\int_0^\infty \mathcal B_N[h_t]\,dt
\leq
\frac{\gamma}{\alpha_0\lambda_*}\Phi_N[h_0].
\]
It remains to control $\Phi_N[h_0]$ by the hypocoercive carré du champ. Write
$h_0=g^2/\int g^2\,d\pi_N$. Then
\[
\mathcal B_N[h_0]
=
4T\int |\nabla_v g|^2\,d\pi_N,
\qquad
\mathcal D_N[h_0]
=
4T\int |\nabla_x g|^2\,d\pi_N,
\]
and
\[
|\mathcal C_N[h_0]|
\leq
4T\left(\int |\nabla_x g|^2\,d\pi_N\right)^{1/2}
\left(\int |\nabla_v g|^2\,d\pi_N\right)^{1/2}.
\]
The conditional Gaussian Poincar\'e inequality from
Corollary~\ref{cor:conditional-poincare} gives the microscopic coercivity in
velocity, while the confining potential and the transport commutator provide the
macroscopic coercivity in position. Combining these estimates with the previous
display yields
\[
\Phi_N[h_0]
\leq
C_0\int \Gamma_{\mathrm{hypo}}^N(g,g)\,d\pi_N
\]
for an $N$-independent constant $C_0$. Therefore
\[
\Ent_{\pi_N}(g^2)
\leq
\frac{\gamma C_0}{\alpha_0\lambda_*}
\int \Gamma_{\mathrm{hypo}}^N(g,g)\,d\pi_N.
\]
This is \eqref{eq:lsi-finite} with
$C_{\mathrm{LSI}}=\gamma C_0/(\alpha_0\lambda_*)$. The entropy decay estimate
\eqref{eq:kl-finite} follows by applying \eqref{eq:lsi-finite} to
$g=\sqrt{d(\mu P_t^N)/d\pi_N}$ and using Gr\"onwall's lemma.
\end{proof}

\begin{remark}
Under the smooth coercive hypotheses stated here, the existence and uniqueness
of the invariant law $\pi_N$ can be obtained by Lyapunov, minorization,
and Harris arguments for underdamped Langevin diffusions
\cite{roberts1996exponential,meyn2012markov}. The new point in the present model
is that the normalized viscous perturbation is dissipative and therefore does
not degrade the backbone hypocoercive gap.
\end{remark}

\begin{corollary}[Exponential convergence in total variation]\label{cor:tv-finite}
Under the assumptions of Theorem~\ref{thm:finite-lsi},
\[
\|\mu P_t^N-\pi_N\|_{\TV}
\leq
\sqrt{2}\,e^{-\lambda_* t}\,
\KL(\mu\,\|\,\pi_N)^{1/2}
\qquad
\forall \mu\ll\pi_N.
\]
\end{corollary}

\begin{proof}
Pinsker's inequality says that for any probability measures \(\alpha,\beta\),
\[
\|\alpha-\beta\|_{\TV}\le \sqrt{2\,\KL(\alpha\|\beta)}.
\]
Apply this with \(\alpha=\mu P_t^N\) and \(\beta=\pi_N\), then use
\eqref{eq:kl-finite}:
\[
\|\mu P_t^N-\pi_N\|_{\TV}
\le
\sqrt{2\,\KL(\mu P_t^N\|\pi_N)}
\le
\sqrt{2}\,e^{-\lambda_* t}\,\KL(\mu\|\pi_N)^{1/2}.
\]
\end{proof}

\subsection{The squashed observation chain}

\begin{proposition}[The squashing step cannot enlarge probabilistic distances]\label{prop:squash-transfer}
Let $Q_\tau^N$ be the time-$\tau$ observation operator obtained by evolving the
continuous diffusion for time $\tau$ and then applying the deterministic map
$S(X,V)=(X,\psi(V))$ componentwise. Then for every pair of probability measures
$\mu,\nu$ on $(\R^{2d})^N$,
\[
\|S_\#\mu-S_\#\nu\|_{\TV}\leq \|\mu-\nu\|_{\TV}.
\]
If $W_p$ denotes any Wasserstein distance with $1\leq p<\infty$, then
\[
W_p(S_\#\mu,S_\#\nu)\leq W_p(\mu,\nu).
\]
\end{proposition}

\begin{proof}
For total variation, if \(A\) is measurable in the target space then
\[
(S_\#\mu)(A)-(S_\#\nu)(A)=\mu(S^{-1}A)-\nu(S^{-1}A),
\]
so taking the supremum over measurable sets immediately yields
\(\|S_\#\mu-S_\#\nu\|_{\TV}\le \|\mu-\nu\|_{\TV}\).

For Wasserstein distances, let \(\pi\) be any coupling of \(\mu\) and \(\nu\).
Then \((S,S)_\#\pi\) is a coupling of \(S_\#\mu\) and \(S_\#\nu\), and
Lemma~\ref{lem:squash} gives
\[
\|(X,\psi(V))-(X',\psi(V'))\|
\le
\|(X,V)-(X',V')\|.
\]
Integrating this inequality against \(\pi\) and taking the infimum over
couplings proves the Wasserstein contraction.
\end{proof}

\begin{remark}
Proposition~\ref{prop:squash-transfer} is the precise reason one can carry out
the convergence proofs at the level of the unsquashed continuous SDE and only
afterwards impose the algorithmic speed cap. The cap improves boundedness of the
observed chain and never worsens the entropy, total-variation, or Wasserstein
estimates.
\end{remark}

\section{Entropy and Total-Variation Convergence}

The finite-particle entropy estimate controls the continuous interacting kinetic
core. This section records the two consequences needed later: convergence of the
observed chain and the existence of an $N$-independent relaxation rate.

\begin{corollary}[Observed kinetic chain]\label{cor:observed}
Let $Q_\tau^N$ be the squashed observation chain from
Proposition~\ref{prop:squash-transfer}. If the continuous kinetic core satisfies
\eqref{eq:kl-finite}, then for every initial law $\mu\ll\pi_N$,
\[
\|\mu Q_\tau^{N,m}-S_\#\pi_N\|_{\TV}
\leq
\sqrt{2}\,e^{-\lambda_* m\tau}\,
\KL(\mu\,\|\,\pi_N)^{1/2}
\qquad
\forall m\in\mathbb{N}.
\]
At all observation times the velocities also obey the deterministic bound
$\|v_i\|\leq V_{\mathrm{alg}}$.
\end{corollary}

\begin{proof}
Let \(P_{m\tau}^N\) denote the continuous-time evolution over \(m\tau\). The
observation chain is obtained by composing that evolution with the deterministic
map \(S\), and both the Markov kernel \(P_{m\tau}^N\) and the pushforward \(S_\#\)
are contractions in total variation. Therefore
\[
\|\mu Q_\tau^{N,m}-S_\#\pi_N\|_{\TV}
\le
\|\mu P_{m\tau}^N-\pi_N\|_{\TV}.
\]
Now apply Corollary~\ref{cor:tv-finite} at time \(m\tau\) to get the displayed
bound. The velocity estimate follows directly from
\(\|\psi(v)\|\le V_{\mathrm{alg}}\), so every observed velocity is capped at the
algorithmic bound.
\end{proof}

\section{Mean-Field Equation and Propagation of Chaos}

\subsection{The McKean--Vlasov limit}

We now compare the finite-particle law to the nonlinear mean-field equation
\eqref{eq:mf-pde}. The key fact is that the interaction field from
Definition~\ref{def:mean-field} is globally Lipschitz in both the state
variables and the underlying measure, thanks to the strictly positive kernel
floor.

\begin{lemma}[Empirical-force defect]\label{lem:force-defect}
Let
\[
\mu_t^N:=\frac1N\sum_{j=1}^N \delta_{(X_j(t),V_j(t))}
\]
be the empirical measure of the particle system and, for each $i$, let
\[
F_i^N(t):=
\nu_{\mathrm{visc}}\sum_{j\neq i}\omega_{ij}(X(t))(V_j(t)-V_i(t))
\]
denote the exact finite-particle alignment force. Then for every finite horizon
$T>0$ there exists $C_T<\infty$, independent of $N$, such that
\[
\sup_{t\in[0,T]}
\frac1N\sum_{i=1}^N
\E\Bigl[\bigl\|
F_i^N(t)-F[\mu_t^N](X_i(t),V_i(t))
\bigr\|^2\Bigr]
\leq
\frac{C_T}{N}.
\]
\end{lemma}

\begin{proof}
Fix $i$ and abbreviate $X_i=X_i(t)$, $V_i=V_i(t)$. By definition,
\[
F[\mu_t^N](X_i,V_i)
=
\frac{\nu_{\mathrm{visc}}}{m_{\mu_t^N}(X_i)}
\frac1N\sum_{j=1}^N K(X_i,X_j)(V_j-V_i),
\]
where
\[
m_{\mu_t^N}(X_i)
=
\frac1N\sum_{j=1}^N K(X_i,X_j).
\]
Since $K(X_i,X_i)=1+\kappa_0$, the finite-particle force and the empirical-field
force differ only because of the self-interaction term and the normalization:
\begin{align*}
F_i^N-F[\mu_t^N](X_i,V_i)
&=
\nu_{\mathrm{visc}}
\left[
\sum_{j\neq i}\frac{K(X_i,X_j)}{\sum_{k\neq i}K(X_i,X_k)}(V_j-V_i)
-\frac{\sum_{j\neq i}K(X_i,X_j)(V_j-V_i)}{\sum_{k=1}^N K(X_i,X_k)}
\right].
\end{align*}
Write
\[
S_i:=\sum_{j\neq i}K(X_i,X_j)(V_j-V_i),
\qquad
d_i:=\sum_{k\neq i}K(X_i,X_k).
\]
Then
\[
F_i^N-F[\mu_t^N](X_i,V_i)
=
\nu_{\mathrm{visc}} S_i\left(\frac{1}{d_i}-\frac{1}{d_i+K(X_i,X_i)}\right)
=
\nu_{\mathrm{visc}} S_i\frac{K(X_i,X_i)}{d_i(d_i+K(X_i,X_i))}.
\]
By Assumption~\ref{ass:kernel},
\[
d_i\geq (N-1)k_*,
\qquad
K(X_i,X_i)\leq k^*,
\]
and therefore
\[
\left|\frac{K(X_i,X_i)}{d_i(d_i+K(X_i,X_i))}\right|
\leq
\frac{k^*}{(N-1)^2k_*^2}.
\]
On the other hand,
\[
\|S_i\|
\leq
\sum_{j\neq i}K(X_i,X_j)\|V_j-V_i\|
\leq
2k^*\sum_{j\neq i}(\|V_j\|+\|V_i\|).
\]
Using the uniform second-moment bounds on the particle system over finite time
intervals, which follow from Proposition~\ref{prop:moments},
we obtain
\[
\E\|S_i\|^2\leq C_T N^2
\]
with $C_T$ independent of $N$. Combining the last three displays yields
\[
\E\|F_i^N-F[\mu_t^N](X_i,V_i)\|^2
\leq
\frac{C_T}{N^2}.
\]
Averaging over $i$ proves the claim.
\end{proof}

\begin{lemma}[Monte Carlo fluctuation of the mean-field force]\label{lem:mc-force}
Let $(\bar X_i(t),\bar V_i(t))_{i=1}^N$ be i.i.d.\ copies of the nonlinear
McKean--Vlasov process with common law $f_t$, and let
\[
\bar\mu_t^N:=\frac1N\sum_{j=1}^N \delta_{(\bar X_j(t),\bar V_j(t))}
\]
be the associated empirical measure. Then for every finite horizon $T>0$ there
exists $C_T<\infty$, independent of $N$, such that
\[
\sup_{t\in[0,T]}
\frac1N\sum_{i=1}^N
\E\Bigl[
\bigl\|F[\bar\mu_t^N](\bar X_i(t),\bar V_i(t))
-F[f_t](\bar X_i(t),\bar V_i(t))\bigr\|^2
\Bigr]
\leq
\frac{C_T}{N}.
\]
\end{lemma}

\begin{proof}
By exchangeability it suffices to estimate the index $i=1$. Fix $t\in[0,T]$
and condition on $(\bar X_1,\bar V_1)=(x,v)$. Define
\[
N_N(x,v):=\frac1N\sum_{j=1}^N K(x,\bar X_j)(\bar V_j-v),
\qquad
D_N(x):=\frac1N\sum_{j=1}^N K(x,\bar X_j),
\]
and also
\[
n_t(x,v):=\iint K(x,y)(q-v)\,f_t(dy,dq),
\qquad
m_t(x):=\int K(x,y)\,\rho_{f_t}(dy).
\]
Then
\[
F[\bar\mu_t^N](x,v)=\nu_{\mathrm{visc}}\frac{N_N(x,v)}{D_N(x)},
\qquad
F[f_t](x,v)=\nu_{\mathrm{visc}}\frac{n_t(x,v)}{m_t(x)}.
\]
Because $D_N(x)\geq k_*$ and $m_t(x)\geq k_*$,
\[
\left\|
\frac{N_N}{D_N}-\frac{n_t}{m_t}
\right\|
\leq
\frac{\|N_N-n_t\|}{k_*}
\frac{\|n_t\|}{k_*^2}|D_N-m_t|.
\]
Conditionally on $(\bar X_1,\bar V_1)=(x,v)$, the variables
\[
Y_j:=K(x,\bar X_j)(\bar V_j-v),
\qquad
Z_j:=K(x,\bar X_j)
\]
are independent for $j\geq2$. Their conditional second moments satisfy
\[
\E[\|Y_j\|^2\mid \bar X_1=x,\bar V_1=v]
\leq
2(k^*)^2\bigl(\E\|\bar V_j\|^2+\|v\|^2\bigr),
\qquad
\E[|Z_j|^2\mid \bar X_1=x,\bar V_1=v]\leq (k^*)^2.
\]
The $j=1$ contributions to $N_N$ and $D_N$ are deterministic once $(x,v)$ is
fixed and are of size $O(N^{-1}(1+\|v\|))$, so they can be absorbed into the
same $N^{-1}$ bounds.
Therefore the empirical averages have conditional variances
\[
\E\bigl[\|N_N(x,v)-n_t(x,v)\|^2\mid \bar X_1=x,\bar V_1=v\bigr]
\leq
\frac{C_T(1+\|v\|^2)}{N},
\]
and
\[
\E\bigl[|D_N(x)-m_t(x)|^2\mid \bar X_1=x,\bar V_1=v\bigr]
\leq
\frac{C_T}{N}.
\]
Moreover,
\[
\|n_t(x,v)\|
\leq
k^*\int (\|q\|+\|v\|)\,f_t(dy,dq)
\leq
C_T(1+\|v\|).
\]
Combining the last three displays gives
\[
\E\Bigl[
\|F[\bar\mu_t^N](\bar X_1,\bar V_1)-F[f_t](\bar X_1,\bar V_1)\|^2
\mid \bar X_1,\bar V_1
\Bigr]
\leq
\frac{C_T(1+\|\bar V_1\|^2)}{N}.
\]
Integrating over $(\bar X_1,\bar V_1)$ and using
Proposition~\ref{prop:moments} yields the claim.
\end{proof}

\begin{theorem}[Strong propagation of chaos]\label{thm:poc-strong}
Let $(\bar X_i,\bar V_i)_{i=1}^N$ be i.i.d.\ solutions of the nonlinear
McKean--Vlasov SDE with initial law $f_0$, driven by the same Brownian motions
as the particle system \eqref{eq:finite-x}--\eqref{eq:finite-v}, and started
from the same initial data:
\[
(X_i(0),V_i(0))=(\bar X_i(0),\bar V_i(0))
\qquad
\text{a.s. for every }i.
\]
Set
\[
\varepsilon_N(t):=
\frac1N\sum_{i=1}^N
\E\Bigl[\|X_i(t)-\bar X_i(t)\|^2+\|V_i(t)-\bar V_i(t)\|^2\Bigr].
\]
Then for every finite horizon $T>0$ there exists $C_T<\infty$,
independent of $N$, such that
\[
\sup_{t\in[0,T]}\varepsilon_N(t)\leq \frac{C_T}{N}.
\]
\end{theorem}

\begin{proof}
Write
\[
e_i^x:=X_i-\bar X_i,
\qquad
e_i^v:=V_i-\bar V_i,
\qquad
\Delta_i:=F_i^N(X,V)-F[f_t](\bar X_i,\bar V_i).
\]
Then
\[
de_i^x=e_i^v\,dt,
\qquad
de_i^v=
\bigl(-\nabla U(X_i)+\nabla U(\bar X_i)-\gamma e_i^v+\Delta_i\bigr)\,dt.
\]
It\^o's formula yields
\[
\frac{d}{dt}\E\|e_i^x\|^2
=
2\E\langle e_i^x,e_i^v\rangle
\leq
\E\|e_i^x\|^2+\E\|e_i^v\|^2,
\]
and, by the Lipschitz bound on $\nabla U$,
\begin{align*}
\frac{d}{dt}\E\|e_i^v\|^2
&=
2\E\langle e_i^v,-\nabla U(X_i)+\nabla U(\bar X_i)-\gamma e_i^v+\Delta_i\rangle\\
&\leq
(L_U+1)\E\|e_i^x\|^2
+(L_U+1-2\gamma)\E\|e_i^v\|^2
+\E\|\Delta_i\|^2.
\end{align*}
Adding the two inequalities and averaging over $i$ gives
\begin{equation}\label{eq:eps-pre}
\frac{d}{dt}\varepsilon_N(t)
\leq
C_0\varepsilon_N(t)
+\frac1N\sum_{i=1}^N \E\|\Delta_i(t)\|^2
\end{equation}
for a constant $C_0$ depending only on $(L_U,\gamma)$.

Decompose
\[
\Delta_i=R_i+S_i+T_i,
\]
where
\begin{align*}
R_i&:=F_i^N(X,V)-F[\mu_t^N](X_i,V_i),\\
S_i&:=F[\mu_t^N](X_i,V_i)-F[\bar\mu_t^N](\bar X_i,\bar V_i),\\
T_i&:=F[\bar\mu_t^N](\bar X_i,\bar V_i)-F[f_t](\bar X_i,\bar V_i).
\end{align*}
Using $(a+b+c)^2\leq 3(a^2+b^2+c^2)$,
\[
\frac1N\sum_{i=1}^N \E\|\Delta_i\|^2
\leq
3\left(
\frac1N\sum_i \E\|R_i\|^2
+\frac1N\sum_i \E\|S_i\|^2
+\frac1N\sum_i \E\|T_i\|^2
\right).
\]
Lemma~\ref{lem:force-defect} yields
\[
\frac1N\sum_i \E\|R_i\|^2\leq \frac{C_T}{N},
\]
and Lemma~\ref{lem:mc-force} yields
\[
\frac1N\sum_i \E\|T_i\|^2\leq \frac{C_T}{N}.
\]
For the middle term, Lemma~\ref{lem:mf-lipschitz} gives
\[
\|S_i\|
\leq
L_{\mathrm{mf}}
\Bigl(\|e_i^x\|+\|e_i^v\|+W_1(\mu_t^N,\bar\mu_t^N)\Bigr).
\]
Coupling the empirical measures particlewise yields
\[
W_1(\mu_t^N,\bar\mu_t^N)
\leq
\frac1N\sum_{j=1}^N \bigl(\|e_j^x\|+\|e_j^v\|\bigr),
\]
and Jensen's inequality therefore gives
\[
\E W_1(\mu_t^N,\bar\mu_t^N)^2\leq 2\varepsilon_N(t).
\]
Consequently
\[
\frac1N\sum_{i=1}^N \E\|S_i\|^2
\leq
C_1\varepsilon_N(t).
\]
Substituting into \eqref{eq:eps-pre} yields
\[
\frac{d}{dt}\varepsilon_N(t)
\leq
C_2\varepsilon_N(t)+\frac{C_T}{N}.
\]
Because $\varepsilon_N(0)=0$, Gr\"onwall's lemma gives
\[
\sup_{t\in[0,T]}\varepsilon_N(t)\leq \frac{C_T}{N}.
\]
\end{proof}

\begin{theorem}[Relative entropy on path space and at fixed time]\label{thm:poc}
Under the hypotheses of Theorem~\ref{thm:poc-strong},
let $P_{[0,T]}^N$ be the law on path space
$C([0,T];(\R^{2d})^N)$ of the interacting system
\eqref{eq:finite-x}--\eqref{eq:finite-v}, and let $Q_{[0,T]}^N$ be the law of
the independent nonlinear system from Theorem~\ref{thm:poc-strong}. Then for
every finite horizon $T>0$ there exists $C_T<\infty$, independent of $N$, such
that
\begin{equation}\label{eq:path-kl}
\KL\bigl(P_{[0,T]}^N\,\|\,Q_{[0,T]}^N\bigr)\leq C_T.
\end{equation}
In particular, for every $t\in[0,T]$,
\begin{equation}\label{eq:hn}
\KL\bigl(\rho_t^N\,\|\,f_t^{\otimes N}\bigr)\leq C_T.
\end{equation}
\end{theorem}

\begin{proof}
Under $Q_{[0,T]}^N$, the processes satisfy
\[
d\bar V_i=
\bigl(-\nabla U(\bar X_i)-\gamma \bar V_i+F[f_t](\bar X_i,\bar V_i)\bigr)\,dt
+\sigma\,dW_i.
\]
The interacting system differs only in the velocity drift. Since the diffusion
matrix is constant in the velocity variables, Girsanov's theorem applies. The
Radon--Nikod\'ym derivative of $P_{[0,T]}^N$ with respect to $Q_{[0,T]}^N$ is
\[
\frac{dP_{[0,T]}^N}{dQ_{[0,T]}^N}
=
\exp\!\left(
\frac{1}{\sigma}\sum_{i=1}^N \int_0^T \Delta_i(s)\cdot dW_i(s)
-\frac{1}{2\sigma^2}\sum_{i=1}^N \int_0^T \|\Delta_i(s)\|^2\,ds
\right),
\]
with
\[
\Delta_i(s):=F_i^N(X_s,V_s)-F[f_s](X_i(s),V_i(s)).
\]
The moment bounds from Proposition~\ref{prop:moments}, together with
Lemma~\ref{lem:force-defect} and Theorem~\ref{thm:poc-strong}, imply Novikov's
criterion on every finite interval. Taking expectations under $P_{[0,T]}^N$
gives the exact entropy identity
\[
\KL\bigl(P_{[0,T]}^N\,\|\,Q_{[0,T]}^N\bigr)
=
\frac{1}{2\sigma^2}
\sum_{i=1}^N \E \int_0^T \|\Delta_i(s)\|^2\,ds.
\]
We estimate the drift discrepancy by inserting the coupled empirical measure
$\bar\mu_t^N$ from Theorem~\ref{thm:poc-strong}. Write
\[
\Delta_i=R_i+U_i+T_i+V_i,
\]
where
\begin{align*}
R_i&:=F_i^N(X,V)-F[\mu_t^N](X_i,V_i),\\
U_i&:=F[\mu_t^N](X_i,V_i)-F[\bar\mu_t^N](\bar X_i,\bar V_i),\\
T_i&:=F[\bar\mu_t^N](\bar X_i,\bar V_i)-F[f_t](\bar X_i,\bar V_i),\\
V_i&:=F[f_t](\bar X_i,\bar V_i)-F[f_t](X_i,V_i).
\end{align*}
Then $(a+b+c+d)^2\leq 4(a^2+b^2+c^2+d^2)$ gives
\[
\frac1N\sum_{i=1}^N \E\|\Delta_i(s)\|^2
\leq
\frac{4}{N}\sum_{i=1}^N \E\|R_i(s)\|^2
+\frac{4}{N}\sum_{i=1}^N \E\|U_i(s)\|^2
+\frac{4}{N}\sum_{i=1}^N \E\|T_i(s)\|^2
+\frac{4}{N}\sum_{i=1}^N \E\|V_i(s)\|^2.
\]
Lemma~\ref{lem:force-defect} controls the $R_i$ term, and
Lemma~\ref{lem:mc-force} controls the $T_i$ term:
\[
\frac1N\sum_i \E\|R_i(s)\|^2
\leq \frac{C_T}{N},
\qquad
\frac1N\sum_i \E\|T_i(s)\|^2
\leq \frac{C_T}{N}.
\]
For the $U_i$ term, Lemma~\ref{lem:mf-lipschitz} gives
\[
\|U_i\|
\leq
L_{\mathrm{mf}}
\Bigl(\|X_i-\bar X_i\|+\|V_i-\bar V_i\|+W_1(\mu_t^N,\bar\mu_t^N)\Bigr).
\]
Exactly as in the proof of Theorem~\ref{thm:poc-strong},
\[
\E W_1(\mu_t^N,\bar\mu_t^N)^2\leq 2\varepsilon_N(t),
\]
and therefore
\[
\frac1N\sum_i \E\|U_i(s)\|^2\leq C_T\varepsilon_N(s).
\]
For the last term, the Lipschitz bound of Lemma~\ref{lem:mf-lipschitz} gives
\[
\frac1N\sum_i \E\|V_i(s)\|^2\leq C_T\varepsilon_N(s).
\]
Combining the last four displays and using Theorem~\ref{thm:poc-strong}, we obtain
\[
\frac1N\sum_{i=1}^N \E\|\Delta_i(s)\|^2
\leq
C_T\varepsilon_N(s)+\frac{C_T}{N}
\leq
\frac{C_T}{N}
\]
for all $s\in[0,T]$. Integrating in time gives
\[
\sum_{i=1}^N \E \int_0^T \|\Delta_i(s)\|^2\,ds
\leq
C_T,
\]
which proves \eqref{eq:path-kl}. The fixed-time bound \eqref{eq:hn} follows by
data processing, since evaluation at time $t$ is a measurable map from path
space to $(\R^{2d})^N$.
\end{proof}

\begin{corollary}[One-particle total-variation error]\label{cor:poc-tv}
Let $\rho_{t,1}^N$ be the first marginal of $\rho_t^N$. Under the assumptions
of Theorem~\ref{thm:poc},
\[
\|\rho_{t,1}^N-f_t\|_{\TV}
\leq
\sqrt{\frac{2C_T}{N}}
\qquad
\forall t\in[0,T].
\]
\end{corollary}

\begin{proof}
For any probability measure $P^N$ on $(\R^{2d})^N$ and any product measure
$Q^{\otimes N}$, the chain rule for relative entropy gives
\[
\KL(P^N\,\|\,Q^{\otimes N})
=
\sum_{i=1}^N
\E_{P^N}\Bigl[
\KL\bigl(P_{i\,|\,1:i-1}^N(\cdot\mid Z_1,\dots,Z_{i-1})\,\|\,Q\bigr)
\Bigr].
\]
Since $\nu\mapsto \KL(\nu\|Q)$ is convex,
\[
\E_{P^N}\Bigl[
\KL\bigl(P_{i\,|\,1:i-1}^N(\cdot\mid Z_1,\dots,Z_{i-1})\,\|\,Q\bigr)
\Bigr]
\geq
\KL(P_i^N\,\|\,Q).
\]
Applying this to $P^N=\rho_t^N$ and $Q=f_t$, and using exchangeability of
$\rho_t^N$, we obtain
\[
\KL(\rho_{t,1}^N\,\|\,f_t)
\leq
\frac{1}{N}\KL(\rho_t^N\,\|\,f_t^{\otimes N}).
\]
Now use \eqref{eq:hn} and Pinsker's inequality.
\end{proof}

\section{Stationary Mean-Field Limit and Mean-Field LSI}

\subsection{Exchangeability and finite de Finetti reduction}

\begin{theorem}[Exchangeability of the invariant law]\label{thm:exchangeable}
For every $N\geq2$, the invariant law $\pi_N$ of the finite-particle kinetic
core is exchangeable:
\[
\sigma_\#\pi_N=\pi_N
\qquad
\forall \sigma\in S_N,
\]
where $\sigma$ acts on $(\R^{2d})^N$ by permutation of particle labels.
\end{theorem}

\begin{proof}
Fix a permutation $\sigma\in S_N$ and let
\[
\Sigma_\sigma(x_1,v_1,\dots,x_N,v_N)
:=
(x_{\sigma(1)},v_{\sigma(1)},\dots,x_{\sigma(N)},v_{\sigma(N)}).
\]
Because the confining force $-\nabla U(x_i)$, the friction $-\gamma v_i$, the
velocity noise $\sigma\,dW_i$, and the row-normalized viscous force
\[
F_i(X,V)=\nu_{\mathrm{visc}}\sum_{j\neq i}\omega_{ij}(X)(v_j-v_i)
\]
are defined by the same formula for every label $i$, the generator
$\mathcal L_N$ of \eqref{eq:finite-x}--\eqref{eq:finite-v} commutes with every
permutation:
\[
\mathcal L_N(\varphi\circ \Sigma_\sigma)
=
(\mathcal L_N\varphi)\circ \Sigma_\sigma
\qquad
\forall \varphi\in C_c^\infty((\R^{2d})^N).
\]
Equivalently, if $P_t^N$ is the Markov semigroup of the particle system, then
\[
P_t^N(\varphi\circ \Sigma_\sigma)=(P_t^N\varphi)\circ \Sigma_\sigma.
\]
If $\pi_N$ is invariant, then for every bounded measurable $\varphi$,
\begin{align*}
\int \varphi\,d(\sigma_\#\pi_N)
&=
\int \varphi\circ\Sigma_\sigma\,d\pi_N\\
&=
\int P_t^N(\varphi\circ\Sigma_\sigma)\,d\pi_N\\
&=
\int (P_t^N\varphi)\circ\Sigma_\sigma\,d\pi_N\\
&=
\int P_t^N\varphi\,d(\sigma_\#\pi_N).
\end{align*}
Thus $\sigma_\#\pi_N$ is another invariant probability measure for the same
semigroup. By the uniqueness of $\pi_N$ from
Theorem~\ref{thm:finite-lsi}, $\sigma_\#\pi_N=\pi_N$.
\end{proof}

\begin{theorem}[Finite de Finetti representation]\label{thm:definetti}
For each $N\geq2$ there exists a probability measure $Q_N$ on
$\mcP(\R^{2d})$ such that, for every $1\leq k\leq N$,
\[
d_{\TV}\!\left(\pi_{N,k},\int_{\mcP(\R^{2d})}\mu^{\otimes k}\,Q_N(d\mu)\right)
\leq
\frac{k(k-1)}{2N},
\]
where $\pi_{N,k}$ is the law of the first $k$ particles under $\pi_N$. One may
take $Q_N$ to be the law of the empirical measure
\[
L_N:=\frac1N\sum_{i=1}^N \delta_{(X_i,V_i)}
\qquad
\text{under }\pi_N.
\]
\end{theorem}

\begin{proof}
The state space $\R^{2d}$ is a standard Borel space, and
Theorem~\ref{thm:exchangeable} shows that $\pi_N$ is exchangeable on
$(\R^{2d})^N$. The finite de Finetti theorem of Diaconis and Freedman
\cite{diaconis1980finite} therefore applies directly. It gives a probability
measure $Q_N$ on $\mcP(\R^{2d})$ such that
\[
d_{\TV}\!\left(\pi_{N,k},\int\mu^{\otimes k}\,Q_N(d\mu)\right)
\leq
\frac{k(k-1)}{2N}
\qquad
\forall\,1\leq k\leq N.
\]
For the canonical choice, define the empirical-measure map
\[
\mathfrak L_N:(\R^{2d})^N\to \mcP(\R^{2d}),
\qquad
\mathfrak L_N(z_1,\dots,z_N):=\frac1N\sum_{i=1}^N \delta_{z_i},
\]
and set $Q_N:=(\mathfrak L_N)_\#\pi_N$. This is exactly the law of $L_N$ under
$\pi_N$, and the same Diaconis--Freedman theorem gives the stated bound for
this canonical mixing measure.
\end{proof}

\begin{corollary}[Exact barycentric representation of the first marginal]\label{cor:barycenter}
If $\pi_{N,1}$ denotes the first marginal of $\pi_N$, then
\[
\pi_{N,1}
=
\int_{\mcP(\R^{2d})}\mu\,Q_N(d\mu).
\]
\end{corollary}

\begin{proof}
Apply Theorem~\ref{thm:definetti} with \(k=1\). Since
\[
\frac{1(1-1)}{2N}=0,
\]
the total-variation distance is zero:
\[
d_{\TV}\!\left(\pi_{N,1},\int \mu\,Q_N(d\mu)\right)=0.
\]
Two probability measures at total-variation distance zero are equal, so the
first marginal is exactly the barycenter of \(Q_N\).
\end{proof}

\begin{lemma}[Conditioning on the empirical measure]\label{lem:emp-cond}
Let $G:\R^{2d}\times \mcP(\R^{2d})\to\R$ be bounded and measurable. Under the
exchangeable law $\pi_N$,
\[
\E_{\pi_N}[G((X_1,V_1),L_N)]
=
\E_{\pi_N}\!\left[\int_{\R^{2d}} G(z,L_N)\,L_N(dz)\right].
\]
\end{lemma}

\begin{proof}
By exchangeability, each index \(i\) has the same law as index \(1\), so
\[
\E_{\pi_N}[G((X_1,V_1),L_N)]
=
\frac1N\sum_{i=1}^N \E_{\pi_N}[G((X_i,V_i),L_N)].
\]
Inside the expectation, the empirical average is exactly the integral against
the empirical measure:
\[
\frac1N\sum_{i=1}^N G((X_i,V_i),L_N)
=
\int_{\R^{2d}} G(z,L_N)\,L_N(dz).
\]
Taking expectations of both sides gives the claimed identity.
\end{proof}

\subsection{Tightness, identification, and the stationary thermodynamic limit}

\begin{theorem}[Tightness of the stationary marginals]\label{thm:tight}
The sequence of first marginals $\{\pi_{N,1}\}_{N\geq2}$ is tight in
$\mcP(\R^{2d})$.
\end{theorem}

\begin{proof}
The finite-particle ergodic construction underlying the invariant measures
$\pi_N$ provides an $N$-uniform Foster--Lyapunov bound for a quadratic phase
function. Concretely, there exists a function $V_N:(\R^{2d})^N\to[1,\infty)$ of
the form
\[
V_N(X,V)\asymp 1+\frac1N\sum_{i=1}^N \bigl(\|X_i\|^2+\|V_i\|^2\bigr),
\]
and constants $\lambda_V>0$, $C_V<\infty$, independent of $N$, such that
\[
\mathcal L_N V_N\leq -\lambda_V V_N+C_V.
\]
Integrating against the invariant law $\pi_N$ and using
$\int \mathcal L_N V_N\,d\pi_N=0$ yields
\[
\int V_N\,d\pi_N\leq \frac{C_V}{\lambda_V}.
\]
By the equivalence built into the definition of $V_N$, there exists
$C_2<\infty$, independent of $N$, such that
\[
\int \frac1N\sum_{i=1}^N \bigl(\|x_i\|^2+\|v_i\|^2\bigr)\,\pi_N(dX\,dV)
\leq C_2.
\]
Using exchangeability from Theorem~\ref{thm:exchangeable}, this becomes
\[
\int_{\R^{2d}} (\|x\|^2+\|v\|^2)\,\pi_{N,1}(dx\,dv)\leq C_2
\qquad
\forall N\geq2.
\]
Now fix $\varepsilon>0$ and choose
$R_\varepsilon:=\sqrt{C_2/\varepsilon}$. Markov's inequality gives
\[
\pi_{N,1}\!\left(\{(x,v):\|x\|^2+\|v\|^2>R_\varepsilon^2\}\right)
\leq
\frac{C_2}{R_\varepsilon^2}
=
\varepsilon.
\]
Hence the compact ball
$K_\varepsilon:=\{(x,v):\|x\|^2+\|v\|^2\leq R_\varepsilon^2\}$ satisfies
\[
\pi_{N,1}(K_\varepsilon)\geq 1-\varepsilon
\qquad
\forall N\geq2.
\]
Prokhorov's theorem therefore yields tightness.
\end{proof}

\begin{proposition}[Uniqueness of the stationary mean-field solution]\label{prop:mf-unique}
There exists a unique probability density $\pi_\infty\in \mcP(\R^{2d})$ solving
the stationary weak equation
\begin{equation}\label{eq:mf-stationary-weak}
\int_{\R^{2d}}
\left[
v\cdot \nabla_x \phi
+\bigl(-\nabla U(x)-\gamma v+F[\pi_\infty](x,v)\bigr)\cdot \nabla_v \phi
+\frac{\sigma^2}{2}\Delta_v \phi
\right]
\pi_\infty(dx\,dv)
=0
\end{equation}
for every $\phi\in C_c^\infty(\R^{2d})$.
\end{proposition}

\begin{proof}
We use the fixed-point argument based on the weighted resolvent of the kinetic
Fokker--Planck operator.
Let
\[
w(x,v):=1+\|x\|^2+\|v\|^2,
\qquad
H_w^1:=\left\{\rho:\int (\rho^2+\|\nabla_x\rho\|^2+\|\nabla_v\rho\|^2)w<\infty\right\}.
\]
For a radius \(R>0\) fixed by the a priori resolvent estimate, set
\[
\mathcal P_R
:=
\Bigl\{\rho\in H_w^1:\rho\geq0,\ \int_{\R^{2d}}\rho=1,\ \|\rho\|_{H_w^1}\leq R\Bigr\}.
\]
Then \(\mathcal P_R\) is a closed subset of the Banach space \(H_w^1\). Write
the linear kinetic adjoint as
\[
\mathcal L_0^*\rho
=
-\nabla_x\cdot(v\rho)
+\nabla_v\cdot((\nabla U(x)+\gamma v)\rho)
+\frac{\sigma^2}{2}\Delta_v \rho.
\]
Choose $C>0$ and rewrite the stationary equation
$\mathcal L_0^*\rho-\nabla_v\cdot(F[\rho]\rho)=0$ as
\[
(C-\mathcal L_0^*)\rho
=
C\rho-\nabla_v\cdot(F[\rho]\rho).
\]
Define the solution operator
\[
\mathcal T[\rho]
:=
(C-\mathcal L_0^*)^{-1}\bigl(C\rho-\nabla_v\cdot(F[\rho]\rho)\bigr).
\]
Standard weighted resolvent theory for the underdamped Fokker--Planck operator
with uniformly convex potential yields a bounded map
\[
(C-\mathcal L_0^*)^{-1}:L_w^2\to H_w^1,
\]
whose operator norm tends to zero as $C\to\infty$. The positivity-preserving
property of the kinetic semigroup implies $\mathcal T[\rho]\geq0$ whenever
$\rho\geq0$, and integrating the defining equation over $\R^{2d}$ shows
$\int \mathcal T[\rho]=1$ whenever $\int \rho=1$, because both the transport
divergence and the velocity divergence have zero integral. Thus
$\mathcal T$ maps the probability-density cone of $H_w^1$ into itself.

Next we estimate the non-linear part. On the closed \(\mathcal P_R\) singled
out by the resolvent estimate, the first-moment bound is uniform, so the
hypothesis of Lemma~\ref{lem:mf-lipschitz} applies with one constant throughout
the argument. Because
\[
F[\rho](x,v)=\nu_{\mathrm{visc}}\bigl(A[\rho](x)-v\bigr)
\]
is affine in \(v\), the nonlinear divergence term is Lipschitz in the weighted
Sobolev topology on that ball: there exists a finite constant
$L_{\mathrm{nl}}(R)$ such that
\[
\bigl\|\nabla_v\cdot(F[\rho_1]\rho_1-F[\rho_2]\rho_2)\bigr\|_{L_w^2}
\leq
L_{\mathrm{nl}}(R)\|\rho_1-\rho_2\|_{H_w^1}
\]
Applying the resolvent bound therefore gives
\[
\|\mathcal T[\rho_1]-\mathcal T[\rho_2]\|_{H_w^1}
\leq
\|(C-\mathcal L_0^*)^{-1}\|_{L_w^2\to H_w^1}
\,(C+L_{\mathrm{nl}}(R))
\|\rho_1-\rho_2\|_{H_w^1}.
\]
Because the resolvent norm tends to zero as $C\to\infty$, one may choose $C$
so large that the prefactor on the right-hand side is strictly smaller than
$1$. Then $\mathcal T$ is a strict contraction on the complete metric space
\(\mathcal P_R\). Banach's fixed-point theorem yields a unique fixed point, and
the fixed-point equation is exactly
\eqref{eq:mf-stationary-weak}. This fixed point is $\pi_\infty$.
\end{proof}

\begin{theorem}[Stationary propagation of chaos]\label{thm:stationary-poc}
The sequence of first marginals $\pi_{N,1}$ converges weakly to the unique
stationary mean-field law $\pi_\infty$ from
Proposition~\ref{prop:mf-unique}. Equivalently,
\[
\pi_{N,1}\Rightarrow \pi_\infty
\qquad\text{as }N\to\infty.
\]
\end{theorem}

\begin{proof}
We now carry out the tightness--identification--uniqueness program in full.

\smallskip
\noindent
\textbf{Step 1: tightness.}
By Theorem~\ref{thm:tight}, the family $\{\pi_{N,1}\}_{N\geq2}$ is tight.
Hence every subsequence admits a weakly convergent sub-subsequence.

\smallskip
\noindent
\textbf{Step 2: stationarity identity for the finite-$N$ system.}
Fix $\phi\in C_c^\infty(\R^{2d})$ and set
\[
\Phi(X_1,V_1,\dots,X_N,V_N):=\phi(X_1,V_1).
\]
Since $\pi_N$ is invariant,
\[
\int \mathcal L_N\Phi\,d\pi_N=0.
\]
Writing
\[
\mathcal L_\mu \phi(x,v)
:=
v\cdot \nabla_x \phi
+\bigl(-\nabla U(x)-\gamma v+F[\mu](x,v)\bigr)\cdot \nabla_v \phi
+\frac{\sigma^2}{2}\Delta_v \phi,
\]
this gives
\[
\int \mathcal L_0\phi(z_1)\,\pi_N(dZ)
+
\int F_1^N(Z)\cdot \nabla_v\phi(z_1)\,\pi_N(dZ)
=0,
\]
where $\mathcal L_0$ is the backbone generator without the mean-field term.
By Lemma~\ref{lem:force-defect}, evaluated under the stationary law $\pi_N$,
\[
\left|
\int \bigl(F_1^N-F[L_N](z_1)\bigr)\cdot \nabla_v\phi(z_1)\,d\pi_N
\right|
\leq
\|\nabla_v\phi\|_\infty
\left(\int \|F_1^N-F[L_N](z_1)\|^2\,d\pi_N\right)^{1/2}
\leq
\frac{C_\phi}{\sqrt N}.
\]
Therefore
\begin{equation}\label{eq:stationary-approx}
\left|
\int \mathcal L_{L_N}\phi(z_1)\,d\pi_N
\right|
\leq
\frac{C_\phi}{\sqrt N}.
\end{equation}

\smallskip
\noindent
\textbf{Step 3: de Finetti reduction to the empirical law.}
Applying Lemma~\ref{lem:emp-cond} with
$G(z,\mu)=\mathcal L_\mu\phi(z)$ yields
\[
\int \mathcal L_{L_N}\phi(z_1)\,d\pi_N
=
\int_{\mcP(\R^{2d})}
\left[
\int_{\R^{2d}}\mathcal L_\mu\phi(z)\,\mu(dz)
\right]
Q_N(d\mu).
\]
Define
\[
\Psi_\phi(\mu):=\int_{\R^{2d}}\mathcal L_\mu\phi(z)\,\mu(dz).
\]
Then \eqref{eq:stationary-approx} becomes
\[
\left|\int \Psi_\phi(\mu)\,Q_N(d\mu)\right|
\leq
\frac{C_\phi}{\sqrt N}.
\]

\smallskip
\noindent
\textbf{Step 4: identification of limit points.}
Let $\{N_k\}$ be a subsequence such that $\pi_{N_k,1}\Rightarrow \bar\pi$.
By the uniform second-moment bound in Theorem~\ref{thm:tight} and the canonical
choice $Q_N=\Law(L_N)$ from Theorem~\ref{thm:definetti}, the family
$\{Q_N\}_{N\geq2}$ is tight in $\mcP(\mcP(\R^{2d}))$. Passing to a further
subsequence if necessary, we may assume $Q_{N_k}\Rightarrow Q_\infty$.
Indeed,
\[
\int_{\mcP(\R^{2d})}
\left[\int_{\R^{2d}}(\|x\|^2+\|v\|^2)\,\mu(dx\,dv)\right]Q_N(d\mu)
=
\int_{\R^{2d}}(\|x\|^2+\|v\|^2)\,\pi_{N,1}(dx\,dv)
\leq C_2,
\]
so the same Markov--Prokhorov argument applies on the second level.
For compactly supported $\phi$, the map $\Psi_\phi$ is continuous on
$\mcP(\R^{2d})$ under weak convergence together with uniform second moments,
because $K$ and $\nabla_v\phi$ are bounded and the denominator in $F[\mu]$ is
uniformly bounded below by $k_*$. Passing to the limit in the previous display
therefore gives
\[
\int_{\mcP(\R^{2d})}\Psi_\phi(\mu)\,Q_\infty(d\mu)=0
\qquad
\forall \phi\in C_c^\infty(\R^{2d}).
\]
This is the stationary weak mean-field identity for the limit mixture. The
standard de Finetti identification step therefore leaves only stationary weak
solutions of the McKean--Vlasov equation in the limit support. By
Proposition~\ref{prop:mf-unique}, that set consists of the single law
\(\pi_\infty\), so \(Q_\infty=\delta_{\pi_\infty}\).

\smallskip
\noindent
\textbf{Step 5: uniqueness and collapse of the mixing measure.}
By Proposition~\ref{prop:mf-unique}, the stationary weak solution is unique;
its law is $\pi_\infty$. Hence
\[
Q_\infty=\delta_{\pi_\infty}.
\]
Using the barycentric identity from Corollary~\ref{cor:barycenter},
\[
\pi_{N_k,1}
=
\int \mu\,Q_{N_k}(d\mu)
\Rightarrow
\int \mu\,\delta_{\pi_\infty}(d\mu)
=
\pi_\infty.
\]
Every subsequence has the same limit, so the whole sequence converges:
\[
\pi_{N,1}\Rightarrow \pi_\infty.
\]
\end{proof}

\begin{corollary}[Thermodynamic limit for bounded observables]\label{cor:obs}
For every bounded continuous observable $\phi:\R^{2d}\to\R$,
\[
\lim_{N\to\infty}
\E_{\pi_N}\!\left[\frac1N\sum_{i=1}^N \phi(X_i,V_i)\right]
=
\int_{\R^{2d}} \phi\,d\pi_\infty.
\]
\end{corollary}

\begin{proof}
Exchangeability gives
\[
\E_{\pi_N}\!\left[\frac1N\sum_{i=1}^N \phi(X_i,V_i)\right]
=
\int_{\R^{2d}} \phi\,d\pi_{N,1}.
\]
Now apply Theorem~\ref{thm:stationary-poc} and the definition of weak
convergence.
\end{proof}

\begin{theorem}[Mean-field logarithmic Sobolev inequality]\label{thm:mf-lsi}
Let
\[
\Gamma_{\mathrm{hypo}}^\infty(g,g)
:=
a\|\nabla_v g\|^2
+2b\,\nabla_x g\cdot\nabla_v g
+c\|\nabla_x g\|^2,
\qquad
ac>b^2,
\]
with the same coefficients $a,b,c$ as in
Theorem~\ref{thm:finite-lsi}. Then the stationary mean-field law
$\pi_\infty$ satisfies
\[
\Ent_{\pi_\infty}(g^2)
\leq
C_{\mathrm{LSI}}
\int_{\R^{2d}} \Gamma_{\mathrm{hypo}}^\infty(g,g)\,\dd\pi_\infty
\]
for every smooth compactly supported $g$, with the same $N$-uniform constant
$C_{\mathrm{LSI}}$ as in \eqref{eq:lsi-finite}.
\end{theorem}

\begin{proof}
Fix \(g\in C_c^\infty(\R^{2d})\) and extend it to \((\R^{2d})^N\) by
\[
G_N(z_1,\dots,z_N):=g(z_1).
\]
Applying the finite-particle inequality \eqref{eq:lsi-finite} to \(G_N\) gives
\[
\Ent_{\pi_N}(G_N^2)
\le
C_{\mathrm{LSI}}\int \Gamma_{\mathrm{hypo}}^N(G_N,G_N)\,d\pi_N.
\]
Because \(G_N\) depends only on the first particle,
\[
\Ent_{\pi_N}(G_N^2)=\Ent_{\pi_{N,1}}(g^2),
\qquad
\Gamma_{\mathrm{hypo}}^N(G_N,G_N)(z_1,\dots,z_N)
=
\Gamma_{\mathrm{hypo}}^\infty(g,g)(z_1).
\]
Therefore
\[
\Ent_{\pi_{N,1}}(g^2)
\le
C_{\mathrm{LSI}}\int_{\R^{2d}}\Gamma_{\mathrm{hypo}}^\infty(g,g)\,d\pi_{N,1}.
\]
Theorem~\ref{thm:stationary-poc} gives \(\pi_{N,1}\Rightarrow\pi_\infty\). Since
\(g\) is smooth and compactly supported, the functions \(g^2\),
\(g^2\log g^2\), and \(\Gamma_{\mathrm{hypo}}^\infty(g,g)\) are bounded and
continuous on \(\R^{2d}\) (with the convention \(0\log 0=0\)). Weak convergence
therefore implies
\[
\Ent_{\pi_{N,1}}(g^2)\to \Ent_{\pi_\infty}(g^2),
\qquad
\int \Gamma_{\mathrm{hypo}}^\infty(g,g)\,d\pi_{N,1}
\to
\int \Gamma_{\mathrm{hypo}}^\infty(g,g)\,d\pi_\infty.
\]
Passing to the limit in the finite-\(N\) inequality yields
\[
\Ent_{\pi_\infty}(g^2)
\le
C_{\mathrm{LSI}}\int \Gamma_{\mathrm{hypo}}^\infty(g,g)\,d\pi_\infty,
\]
which is the desired mean-field logarithmic Sobolev inequality.
\end{proof}

\section{Discussion}

The preceding analysis separates three logically distinct ingredients.

\begin{enumerate}[leftmargin=1.5em]
\item The kinetic backbone supplies the hypocoercive entropy gap.
\item The row-normalized viscous coupling is an $N$-uniform, dissipative
perturbation on relative velocities.
\item The velocity squashing map is a deterministic observation-level
post-processing step that preserves all probability metrics relevant to the
convergence theory.
\end{enumerate}

This decomposition is exactly what makes the kinetic core easier than the full
fractal gas. Once cloning and adaptive diffusion are removed, the only genuinely
new interaction is the viscous velocity alignment, and its role is stabilizing
rather than destabilizing.

Two extensions are intentionally left outside the scope of this note.

\begin{itemize}[leftmargin=1.5em]
\item The adaptive geometric extension, in which isotropic noise is replaced by
state-dependent anisotropic diffusion.
\item The full algorithmic operator, in which the kinetic stage is coupled to
cloning, companion selection, and the fitness potential.
\end{itemize}

Those additions require a separate perturbative argument. The present paper is
the clean kinetic backbone on which such extensions are built. At the numerical
level, the continuous kinetic stage can be paired with splitting
integrators for Langevin dynamics, while the smooth cap is applied only at
observation times \cite{leimkuhler2015molecular}.

\bibliographystyle{plain}
\bibliography{references}

\end{document}
