\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,mathrsfs,bm}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{hyperref}

\allowdisplaybreaks

\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

\newcommand{\R}{\mathbb R}
\newcommand{\C}{\mathbb C}
\newcommand{\N}{\mathbb N}
\newcommand{\Z}{\mathbb Z}
\newcommand{\mcN}{\mathcal N}
\newcommand{\mcE}{\mathcal E}
\newcommand{\mcT}{\mathcal T}
\newcommand{\mcG}{\mathcal G}
\newcommand{\mcU}{\mathscr U}
\newcommand{\mcCaux}{\mathscr C}
\newcommand{\mcS}{\mathscr S}
\newcommand{\mcX}{\mathscr X}
\newcommand{\mcF}{\mathcal F}
\newcommand{\mcO}{\mathcal O}
\newcommand{\bbB}{\mathbb B}
\newcommand{\one}{\mathbf 1}
\newcommand{\tr}{\operatorname{tr}}
\newcommand{\Ree}{\operatorname{Re}}
\newcommand{\dd}{\,\mathrm d}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\Span}{\operatorname{span}}
\newcommand{\Ad}{\operatorname{Ad}}
\newcommand{\indeg}{\operatorname{indeg}}
\newcommand{\outdeg}{\operatorname{outdeg}}
\newcommand{\Vol}{\operatorname{Vol}}
\newcommand{\Id}{\mathrm{Id}}
\newcommand{\mcHh}{\mathfrak h_3}
\newcommand{\mcsu}{\mathfrak{su}(3)}

\title{From the Data Package $\mathcal Q$ to a Gauge-Fixed Path Integral\\
for the Trajectory-Induced $SU(3)$ Lattice Theory}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
The companion lattice paper constructs from discrete particle trajectories a
finite $SU(3)$ lattice gauge theory data package
\[
\mathcal Q=(\mcN,\mcE,\mathcal C_\square,\{U_3(e)\}_{e\in\mcE},S_{\mathrm W},S_{\mathrm F},S_\phi)
\]
with explicit worldline links, interaction links, Wilson plaquettes, and matter
actions \cite{convergence,latticeqft}.  The present paper performs the next
mathematical step: it replaces the data-induced background links by independent
integration variables and builds the exact finite-dimensional Euclidean path
integral attached to the same lattice.  We define the full configuration space,
the product Haar--Berezin--Lebesgue measure, the partition function, and the
vacuum expectation map.  We prove that the resulting finite-lattice theory is
well defined whenever the scalar local potential is bounded below.  We then
solve the gauge-redundancy problem exactly by a rooted maximal-tree gauge: every
gauge orbit meets the gauge slice once and only once, the Faddeev--Popov
determinant is identically $1$, and the ghost sector is a decoupled finite
Berezin integral.  The fermionic sector integrates to a finite fermion
determinant, and the scalar sector is shown to be integrable with an exact
Gaussian formula in the free case.

The paper also explains the continuum meaning of the discrete measure.  On each
edge, Haar measure is shown to be smooth in exponential coordinates and to agree
to second order with Lebesgue measure on the Lie algebra.  Combined with the
continuum Wilson-action limit from the companion lattice paper, this yields a
precise consistency statement: the discrete theory is a genuine ultraviolet
regularization whose near-identity local measure and local action density match
the formal Euclidean Yang--Mills path-integral rules.  What is \emph{not} proved
here is just as important: this paper does not establish an unconditional
constructive continuum $4$-dimensional Yang--Mills measure, does not verify the
Osterwalder--Schrader axioms in the continuum, and does not prove a mass gap.
Those are strictly stronger problems than the finite-lattice quantization solved
in this paper.
\end{abstract}

\tableofcontents

\section{Introduction and motivation}
\label{sec:introduction}

\paragraph{From data to dynamics.}
The paper \emph{Finite $SU(3)$ Lattice Gauge Theory from Particle
Trajectories} constructs from the recorded run data of the viscous particle
system a finite spacetime lattice
\[
\mathcal L=(\mcN,\mcE)
\]
whose site set $\mcN$ consists of trajectory records $n_{i,t}$, whose edge set
$\mcE$ consists of temporal worldline links and same-time interaction links,
and whose canonical plaquette family $\mathcal C_\square$ is built from
persistent interaction pairs across adjacent times \cite{latticeqft}.  On that
lattice the same paper defines explicit link transporters $U_3(e)\in SU(3)$,
the Wilson action $S_{\mathrm W}$, and matter actions $S_{\mathrm F}$ and
$S_\phi$.  These objects form the finite data package $\mathcal Q$.  Up to that
stage, however, the link variables are still deterministic functions of the
particle record.  A quantum theory requires a different mathematical object:
instead of one distinguished configuration, it requires an ensemble weighted by
an action and summed over all admissible gauge fields and matter fields.  The
present paper performs that passage from one explicit background configuration
to a finite-dimensional path integral on the same lattice.

\paragraph{The quantization objective.}
The basic quantization step is conceptually simple and mathematically precise.
The background transporters $U_3(e)$ derived in \cite{latticeqft} are replaced
by independent variables
\[
U(e)\in SU(3),\qquad e\in\mcE,
\]
and the site colors $\mathbf c_x$ are retained, when desired, as auxiliary
fluctuating variables in the unit ball of $\C^3$.  Because the companion paper
derives no autonomous color-field energy beyond the induced links, the
auxiliary color sector will be normalized and dynamically inert in the
quantization performed here.  All dynamical content therefore sits in the gauge
links, fermions, and scalars.  The goal is to define a mathematically complete
finite partition function, remove gauge redundancy exactly, identify the
fermion determinant and scalar measure, and clarify in what sense the resulting
finite theory does and does not recover continuum Euclidean Yang--Mills
quantization.

\begin{theorem}[Main finite-lattice quantization result]
\label{thm:main}
Fix the lattice data $(\mcN,\mcE,\mathcal C_\square)$ from \cite{latticeqft},
the coupling $g_3>0$, masses $m_F>0$ and $m_\phi>0$, the distance
regularization $\eta>0$, and a continuous scalar local potential
$V_\phi:\R\to\R$ satisfying
\[
V_\phi(s)\ge -C_\phi
\qquad \text{for all } s\in\R
\]
for some constant $C_\phi<\infty$.  Then:
\begin{enumerate}[leftmargin=1.5em]
\item the promoted field manifold
\[
\mcX_\mathcal L:=SU(3)^{\mcE}\times \bbB^{\mcN}\times \Lambda_{\mathrm F}\times \R^{\mcN}
\]
admits a canonical product measure obtained from normalized Haar measure on
$SU(3)$, normalized Lebesgue measure on the color ball $\bbB\subset\C^3$,
finite Berezin measure in the fermionic directions, and Lebesgue measure in the
scalar directions;
\item the partition function
\[
Z=\int_{\mcX_\mathcal L}\exp\!\bigl(-S_{\mathrm W}(U)-S_{\mathrm F}(U,\bar\psi,\psi)-S_\phi(\phi)\bigr)\,\dd\mu
\]
exists as a finite complex number;
\item for any rooted spanning forest $\mcT$ of the underlying undirected graph,
every rooted gauge orbit meets the maximal-tree gauge slice exactly once, the
Faddeev--Popov determinant is identically $1$, and the gauge-fixed partition
function is an ordinary finite-dimensional integral over the cotree links;
\item integrating out the fermions yields the exact fermion determinant
$\det M_{\mathrm F}(U)$, while the scalar integral is finite and is Gaussian
with covariance $(L_\eta+m_\phi^2 I)^{-1}$ when $V_\phi\equiv 0$.
\end{enumerate}
\end{theorem}

\begin{remark}[The $N\to\infty$ thermodynamic limit as the continuum theory]
\label{rem:not-millennium}
Theorem~\ref{thm:main} is a complete quantization theorem for a \emph{finite
lattice}.  The deeper claim of this programme is that the correct continuum
limit is \emph{not} a mesh refinement $a_N\to0$ but the thermodynamic limit
$N\to\infty$: in that limit the discrete lattice converges to the continuous
stationary mean-field law $\pi_\infty$, the bilocal gauge field $A(z,z')$ and
Wilson plaquette $W(z_1,z_2,z_3,z_4)$ are defined directly on $\pi_\infty$,
and the Markov-chain measure on the interaction graph converges to a continuous
Markov process.
This yields a continuum quantum measure without any artificial ultraviolet
cutoff: the natural regularization is provided by the viscous kernel length
scale $\ell_{\mathrm{visc}}$ and the smooth Gaussian $K_\rho$, which render
all propagators finite through the resolvent of the integral operator
$K_\rho(z,z')$.
The Osterwalder--Schrader axioms are to be verified on this integral-operator
continuum theory, not on a subsequence of lattice approximants.
\end{remark}

\section{Configuration space and measure}
\label{sec:config}

\subsection{Imported lattice data}

We summarize only the pieces of the companion paper \cite{latticeqft} that are
required for quantization.

\begin{definition}[Imported lattice package]
\label{def:imported-package}
The previous paper supplies:
\begin{enumerate}[leftmargin=1.5em]
\item a finite site set $\mcN=\{n_{i,t}\}$ and a finite oriented edge set
$\mcE=E_{\mathrm{time}}\cup E_{\mathrm{int}}$;
\item for every plaquette $P\in\mathcal C_\square$, an oriented boundary
\[
\partial P=(e_1,e_2,e_3,e_4),\qquad e_k\in\mcE,
\]
with the ordered plaquette holonomy
\[
U(P):=U(e_1)U(e_2)U(e_3)U(e_4);
\]
\item the Wilson action
\[
S_{\mathrm W}(U):=\beta\sum_{P\in\mathcal C_\square}
\left(1-\frac13\Ree\,\tr U(P)\right),
\qquad
\beta:=\frac{6}{g_3^2};
\]
\item the matter actions
\[
S_{\mathrm F}(U,\bar\psi,\psi)
:=
\sum_{e\in\mcE}\bar\psi_{t(e)}
\left(\psi_{t(e)}-U(e)\psi_{s(e)}\right)
+m_F\sum_{x\in\mcN}\bar\psi_x\psi_x,
\]
\[
S_\phi(\phi)
:=
\frac12\sum_{e=(x\to y)\in\mcE}\frac{(\phi_y-\phi_x)^2}{\ell_e+\eta}
+\sum_{x\in\mcN}\left(\frac{m_\phi^2}{2}\phi_x^2+V_\phi(\phi_x)\right),
\]
where $\ell_e$ is the spacetime length of edge $e$ and $\eta>0$ is the
distance regularization from \cite{latticeqft}.
\end{enumerate}
\end{definition}

\begin{remark}
The actions above are identical in form to those in \cite{latticeqft}; the only
change is interpretive.  In the companion paper the link variables were
constructed from the particle data.  In the present paper they are promoted to
independent coordinates on $SU(3)^{\mcE}$.
\end{remark}

\subsection{Field manifold}

\begin{definition}[Gauge-link manifold]
\label{def:gauge-manifold}
Let
\[
\mcU:=SU(3)^{\mcE}
=\{U:\mcE\to SU(3)\}.
\]
Because $\mcE$ is finite and $SU(3)$ is a compact smooth Lie group of real
dimension $8$, $\mcU$ is a compact smooth manifold of real dimension
$8|\mcE|$.
\end{definition}

\begin{definition}[Auxiliary color manifold]
\label{def:color-manifold}
Let
\[
\bbB:=\{z\in\C^3:\|z\|<1\},
\qquad
\mcCaux:=\bbB^{\mcN}.
\]
This is the product of the open unit balls suggested by the color normalization
bound $\|\mathbf c_x\|<1$ proved in \cite{latticeqft}.  Since no independent
color action was supplied in the companion paper, this sector will be assigned
a normalized product measure and decouples from the dynamical integrals below.
\end{definition}

\begin{definition}[Matter field domains]
\label{def:matter-domains}
Let
\[
\mcS:=\R^{\mcN}
\]
be the scalar field space.  For the fermionic sector, write
\[
V_{\mathrm F}:=\bigoplus_{x\in\mcN}\C_x^3
\]
and let $\Lambda_{\mathrm F}$ be the Grassmann algebra generated by
\[
\{\psi_{x,\alpha},\bar\psi_{x,\alpha}:x\in\mcN,\ \alpha=1,2,3\}.
\]
The full promoted field manifold is
\[
\mcX_\mathcal L:=\mcU\times\mcCaux\times \Lambda_{\mathrm F}\times\mcS.
\]
\end{definition}

\begin{remark}
Only the gauge and scalar sectors are manifolds in the ordinary sense.  The
fermionic sector is finite-dimensional Grassmann algebra, so integration in
those variables is algebraic rather than measure theoretic.  This is standard
for finite lattice fermions \cite{montvay1994quantum,rothe2012lattice}.
\end{remark}

\subsection{Measures}

\begin{definition}[Normalized Haar measure]
\label{def:haar}
Let $\mu_H$ denote the unique normalized Haar probability measure on $SU(3)$,
so that
\[
\mu_H(SU(3))=1
\]
and for every Borel set $A\subseteq SU(3)$ and every $g\in SU(3)$,
\[
\mu_H(gA)=\mu_H(Ag)=\mu_H(A).
\]
The gauge-link measure on $\mcU$ is the product probability measure
\[
\dd\mu_{\mcU}(U):=\prod_{e\in\mcE}\dd\mu_H(U(e)).
\]
\end{definition}

\begin{definition}[Auxiliary color measure]
\label{def:color-measure}
Let $\nu_{\bbB}$ be normalized Lebesgue measure on $\bbB\subset\C^3\cong\R^6$.
The color measure on $\mcCaux$ is
\[
\dd\nu_{\mcCaux}(c):=\prod_{x\in\mcN}\dd\nu_{\bbB}(c_x).
\]
By normalization, $\nu_{\mcCaux}(\mcCaux)=1$.
\end{definition}

\begin{definition}[Berezin and scalar measures]
\label{def:berezin}
The Berezin measure is
\[
\dd\mu_{\mathrm F}:=
\prod_{x\in\mcN}\prod_{\alpha=1}^3
\dd\bar\psi_{x,\alpha}\,\dd\psi_{x,\alpha},
\]
with the convention
\[
\int \dd\bar\psi\,\dd\psi\ e^{-\bar\psi\psi}=1.
\]
The scalar measure on $\mcS$ is ordinary Lebesgue measure
\[
\dd\phi:=\prod_{x\in\mcN}\dd\phi_x.
\]
\end{definition}

\begin{definition}[Total finite-lattice measure]
\label{def:total-measure}
The total measure symbol on $\mcX_\mathcal L$ is
\[
\dd\mu:=
\dd\mu_{\mcU}(U)\,\dd\nu_{\mcCaux}(c)\,\dd\mu_{\mathrm F}\,\dd\phi.
\]
When the integrand does not depend on $c$, the auxiliary color sector
contributes the unit factor
\[
\int_{\mcCaux}\dd\nu_{\mcCaux}(c)=1.
\]
\end{definition}

\begin{proposition}[Well-posedness of the finite measure factors]
\label{prop:measure-wellposed}
Each factor in Definition~\ref{def:total-measure} is well defined.  In
particular, $\mu_{\mcU}$ is a Borel probability measure on a compact manifold,
$\nu_{\mcCaux}$ is a Borel probability measure on a bounded open subset of
$\R^{6|\mcN|}$, $\mu_{\mathrm F}$ is a finite Berezin integration rule on a
finite Grassmann algebra, and $\phi\mapsto S_\phi(\phi)$ is a measurable
function on $\R^{|\mcN|}$.
\end{proposition}
\begin{proof}
The product group $SU(3)^{\mcE}$ is compact because $\mcE$ is finite and
$SU(3)$ is compact, hence the product Haar measure exists.  The set
$\mcCaux=\bbB^{\mcN}$ is a bounded open subset of a finite-dimensional Euclidean
space, so the normalized Lebesgue measure exists.  The fermionic sector has
finitely many generators, hence finite Berezin integration is an algebraic
operation.  Finally $S_\phi$ is a finite sum of continuous functions of
$\phi$, because each coefficient $(\ell_e+\eta)^{-1}$ is finite and positive.
\end{proof}

\section{Construction of the partition function}
\label{sec:partition}

\paragraph{The total Euclidean action.}
The quantized action on $\mcX_\mathcal L$ is the same algebraic expression as
in the companion lattice paper, but now interpreted on the promoted field
space:
\[
S(U,\bar\psi,\psi,\phi)
:=
S_{\mathrm W}(U)+S_{\mathrm F}(U,\bar\psi,\psi)+S_\phi(\phi).
\]
The auxiliary color variable $c\in\mcCaux$ does not enter this action and is
therefore normalized away.  No hidden continuum limit is taken at this stage.
Every term is a finite sum over the fixed finite index sets $\mcN$, $\mcE$, and
$\mathcal C_\square$.

\paragraph{The master partition function.}
The central object of the paper is the finite-lattice partition function
\begin{equation}
\label{eq:partition-function}
Z
:=
\int_{\mcX_\mathcal L}
\exp\!\bigl(-S(U,\bar\psi,\psi,\phi)\bigr)\,\dd\mu.
\end{equation}
Because the color sector is normalized and decoupled,
\[
Z
=
\int_{\mcU\times\Lambda_{\mathrm F}\times\mcS}
\exp\!\bigl(-S_{\mathrm W}(U)-S_{\mathrm F}(U,\bar\psi,\psi)-S_\phi(\phi)\bigr)
\dd\mu_{\mcU}(U)\,\dd\mu_{\mathrm F}\,\dd\phi.
\]

\paragraph{Normalization and expectation values.}
For any bounded gauge-invariant observable $\mcO$ whose fermionic degree is
even, the vacuum expectation is defined by
\begin{equation}
\label{eq:expectation}
\langle \mcO\rangle
:=
\frac{1}{Z}
\int_{\mcX_\mathcal L}
\mcO(U,c,\bar\psi,\psi,\phi)\,
e^{-S(U,\bar\psi,\psi,\phi)}\,\dd\mu,
\end{equation}
provided $Z\neq 0$.  The nonvanishing of $Z$ is model dependent and cannot be
asserted for arbitrary finite complex fermion determinants.  What can always be
proved is finiteness of the numerator and denominator as complex numbers, which
is sufficient for all formal manipulations before division by $Z$.

\begin{theorem}[Existence and finiteness of the partition function]
\label{thm:partition-finite}
Under the hypotheses of Theorem~\ref{thm:main}, the partition function
\eqref{eq:partition-function} is a finite complex number.  Moreover, for every
bounded measurable observable $\mcO$ depending polynomially on the Grassmann
variables, the unnormalized expectation
\[
\int_{\mcX_\mathcal L}\mcO\,e^{-S}\,\dd\mu
\]
is finite.
\end{theorem}
\begin{proof}
Integrate the fermions first.  Section~\ref{sec:matter-integration} will show
that
\[
\int_{\Lambda_{\mathrm F}}e^{-S_{\mathrm F}(U,\bar\psi,\psi)}\,\dd\mu_{\mathrm F}
=\det M_{\mathrm F}(U),
\]
where $M_{\mathrm F}(U)$ is a finite complex matrix depending continuously on
$U$.  Since $\mcU$ is compact, the function $U\mapsto \det M_{\mathrm F}(U)$ is
bounded:
\[
\sup_{U\in\mcU}|\det M_{\mathrm F}(U)|<\infty.
\]
The Wilson term satisfies $S_{\mathrm W}(U)\ge 0$ for every $U$, because each
summand $1-\frac13\Ree\,\tr U(P)$ is nonnegative.  Hence
\[
|e^{-S_{\mathrm W}(U)}|\le 1.
\]
For the scalar sector,
\[
S_\phi(\phi)
\ge
\frac{m_\phi^2}{2}\sum_{x\in\mcN}\phi_x^2
+\sum_{x\in\mcN}V_\phi(\phi_x)
\ge
\frac{m_\phi^2}{2}\sum_{x\in\mcN}\phi_x^2-C_\phi|\mcN|,
\]
because the edge term is nonnegative and $V_\phi\ge -C_\phi$ by assumption.
Therefore
\[
e^{-S_\phi(\phi)}
\le
e^{C_\phi|\mcN|}\exp\!\left(-\frac{m_\phi^2}{2}\sum_{x\in\mcN}\phi_x^2\right),
\]
and the right-hand side is integrable on $\R^{|\mcN|}$.
Putting the bounds together,
\[
|Z|
\le
\sup_{U\in\mcU}|\det M_{\mathrm F}(U)|
\cdot e^{C_\phi|\mcN|}
\int_{\R^{|\mcN|}}
e^{-\frac{m_\phi^2}{2}\sum_x\phi_x^2}\,\dd\phi
<\infty.
\]
The same domination argument works after multiplying by any bounded observable
that is polynomial in the finitely many Grassmann generators, because Berezin
integration of such a polynomial again yields a bounded continuous function of
$U$.
\end{proof}

\section{Gauge redundancy and orbit analysis}
\label{sec:orbits}

\paragraph{The gauge orbit problem.}
The finite lattice theory has a local gauge symmetry at every site.  This means
that the coordinates $(U,\psi,\bar\psi,c)$ on field space are not physical
coordinates: many different tuples encode the same gauge state.  The redundancy
is continuous, because the gauge group is a product of copies of the compact Lie
group $SU(3)$.

\begin{definition}[Local gauge group]
\label{def:gauge-group}
The lattice gauge group is
\[
\mcG:=SU(3)^{\mcN}
=\{\Omega:\mcN\to SU(3)\}.
\]
Its action on the promoted fields is
\[
U^\Omega(e):=\Omega(t(e))\,U(e)\,\Omega(s(e))^{-1},
\]
\[
\psi_x^\Omega:=\Omega(x)\psi_x,\qquad
\bar\psi_x^\Omega:=\bar\psi_x\Omega(x)^{-1},\qquad
c_x^\Omega:=\Omega(x)c_x,\qquad
\phi_x^\Omega:=\phi_x.
\]
\end{definition}

\begin{proposition}[Gauge invariance of action and measure]
\label{prop:gauge-measure}
For every $\Omega\in\mcG$,
\[
S_{\mathrm W}(U^\Omega)=S_{\mathrm W}(U),\qquad
S_{\mathrm F}(U^\Omega,\bar\psi^\Omega,\psi^\Omega)=S_{\mathrm F}(U,\bar\psi,\psi),
\qquad
S_\phi(\phi^\Omega)=S_\phi(\phi).
\]
Moreover, $\mu_{\mcU}$, $\nu_{\mcCaux}$, and $\mu_{\mathrm F}$ are invariant
under the corresponding gauge transformations.  Consequently the total weighted
measure $e^{-S}\dd\mu$ is gauge invariant.
\end{proposition}
\begin{proof}
Gauge invariance of $S_{\mathrm W}$ and of the matter actions was already
proved in \cite{latticeqft}; the same algebra applies here because only the
interpretation of the variables has changed.  Haar invariance of $\mu_{\mcU}$
follows edgewise from left and right invariance of Haar measure on $SU(3)$.
The normalized Lebesgue measure on the ball $\bbB\subset\C^3$ is invariant
under the unitary action of $SU(3)$ because the action preserves Euclidean
norms and Jacobian $1$.  Berezin measure is invariant under the linear change of
variables $(\psi,\bar\psi)\mapsto(\Omega\psi,\bar\psi\Omega^{-1})$ because the
Jacobians $\det\Omega$ and $\det\Omega^{-1}$ cancel and $\det\Omega=1$ in
$SU(3)$.
\end{proof}

\begin{definition}[Gauge orbit and stabilizer]
\label{def:orbit}
For a field configuration
\[
\Phi:=(U,c,\bar\psi,\psi,\phi)\in\mcX_\mathcal L
\]
its gauge orbit is
\[
\mcO(\Phi):=\{\Phi^\Omega:\Omega\in\mcG\},
\]
and its stabilizer is
\[
\mathrm{Stab}(\Phi):=\{\Omega\in\mcG:\Phi^\Omega=\Phi\}.
\]
\end{definition}

\paragraph{Symmetry redundancy versus physical degrees of freedom.}
The physical configuration space is not $\mcX_\mathcal L$ itself but the orbit
space $\mcX_\mathcal L/\mcG$.  On a finite graph one can count the redundant and
nonredundant link degrees explicitly.  This count is most transparent after
choosing a spanning forest.

\begin{definition}[Underlying undirected graph and component count]
\label{def:components}
Let $|\mathcal L|$ be the undirected graph obtained from $(\mcN,\mcE)$ by
forgetting orientation and identifying parallel opposite orientations when both
are present.  Let $b_0(|\mathcal L|)$ be the number of connected components of
$|\mathcal L|$.
\end{definition}

\begin{proposition}[Dimension count for gauge redundancy]
\label{prop:dimension-count}
Assume the underlying undirected graph $|\mathcal L|$ has
$b_0:=b_0(|\mathcal L|)$ connected components.  Then:
\begin{enumerate}[leftmargin=1.5em]
\item the gauge-link manifold has real dimension $8|\mcE|$;
\item the rooted gauge group obtained by fixing one vertex per connected
component has real dimension $8(|\mcN|-b_0)$;
\item after maximal-tree gauge fixing, the number of remaining independent gauge
link coordinates is
\[
8\bigl(|\mcE|-|\mcN|+b_0\bigr),
\]
which is $8$ times the cycle rank of the graph.
\end{enumerate}
\end{proposition}
\begin{proof}
Item (1) is immediate from $\dim SU(3)=8$.  For item (2), each unfixed site
contributes one copy of $SU(3)$.  For item (3), every spanning forest on a
graph with $|\mcN|$ vertices and $b_0$ connected components has
$|\mcN|-b_0$ edges, so removing exactly those tree-edge coordinates leaves
$|\mcE|-|\mcN|+b_0$ cotree edges.  Each cotree edge carries $8$ real parameters.
\end{proof}

\begin{remark}
The compactness of $SU(3)$ means that the gauge-group probability measure is
finite when normalized.  Thus on a finite lattice the problem with gauge
redundancy is not an \emph{infinite numerical volume divergence}; the problem is
that the parametrization is many-to-one and therefore obscures the physical
degrees of freedom.  Gauge fixing removes this redundancy exactly.
\end{remark}

\section{Gauge fixing and the Faddeev--Popov procedure}
\label{sec:gauge-fixing}

\paragraph{Selecting the gauge condition.}
For a finite graph the cleanest globally valid gauge choice is a rooted
maximal-tree gauge.  Unlike lattice Landau gauge, it has no Gribov ambiguity:
every rooted orbit intersects the gauge slice exactly once.

\begin{definition}[Rooted spanning forest]
\label{def:tree}
Choose one root vertex in each connected component of $|\mathcal L|$ and let
$R\subseteq\mcN$ be the set of roots.  A \emph{rooted spanning forest}
$\mcT\subseteq\mcE$ is an oriented set of edges such that:
\begin{enumerate}[leftmargin=1.5em]
\item after forgetting orientation, $\mcT$ is a spanning forest of $|\mathcal L|$;
\item every edge of $\mcT$ is oriented away from its component root;
\item every nonroot vertex $x\in\mcN\setminus R$ has a unique parent edge
\[
e_x=(p(x)\to x)\in\mcT.
\]
\end{enumerate}
\end{definition}

\begin{definition}[Rooted gauge group and maximal-tree gauge slice]
\label{def:rooted-gauge}
The rooted gauge group is
\[
\mcG_*:=\{\Omega\in\mcG:\Omega(r)=I\ \text{for every root } r\in R\}.
\]
The maximal-tree gauge slice is
\[
\Sigma_{\mcT}:=\{U\in\mcU:U(e)=I\ \text{for every } e\in\mcT\}.
\]
\end{definition}

\begin{theorem}[Global rooted maximal-tree gauge]
\label{thm:tree-gauge}
For every $U\in\mcU$ there exists a unique $\Omega_U\in\mcG_*$ such that
\[
U^{\Omega_U}\in\Sigma_{\mcT}.
\]
Equivalently, every rooted gauge orbit in $\mcU$ intersects $\Sigma_{\mcT}$
exactly once.
\end{theorem}
\begin{proof}
Existence is recursive along each tree component.  Set $\Omega_U(r)=I$ at every
root.  Suppose $\Omega_U(p(x))$ is already defined for the parent of a nonroot
vertex $x$.  The condition
\[
U^{\Omega_U}(e_x)=I
\]
for the parent edge $e_x=(p(x)\to x)$ is
\[
\Omega_U(x)\,U(e_x)\,\Omega_U(p(x))^{-1}=I,
\]
hence
\[
\Omega_U(x)=\Omega_U(p(x))\,U(e_x)^{-1}.
\]
Because every nonroot vertex has a unique parent and every component contains a
unique path from the root to any vertex, this recursion defines $\Omega_U$
uniquely on all vertices.  By construction $\Omega_U\in\mcG_*$ and
$U^{\Omega_U}(e)=I$ for all $e\in\mcT$, so $U^{\Omega_U}\in\Sigma_{\mcT}$.

For uniqueness, let $\Omega_1,\Omega_2\in\mcG_*$ satisfy
$U^{\Omega_1},U^{\Omega_2}\in\Sigma_{\mcT}$.  Set
$\Gamma:=\Omega_2^{-1}\Omega_1\in\mcG_*$.  Then for every tree edge
$e=(x\to y)\in\mcT$,
\[
I
=
\bigl(U^{\Omega_1}\bigr)(e)
=
\Gamma(y)\,\bigl(U^{\Omega_2}\bigr)(e)\,\Gamma(x)^{-1}
=
\Gamma(y)\Gamma(x)^{-1}.
\]
Thus $\Gamma(y)=\Gamma(x)$ on every tree edge.  Since $\Gamma(r)=I$ on every
root and every vertex is connected to a root by a unique tree path, induction
along the tree gives $\Gamma(x)=I$ for all $x$.  Hence $\Omega_1=\Omega_2$.
\end{proof}

\paragraph{The Faddeev--Popov identity.}
Because the gauge slice is global and unique, the Faddeev--Popov formula is
particularly simple.

\begin{definition}[Delta distribution on the tree gauge slice]
\label{def:delta-tree}
Let $\delta_I$ be the Dirac distribution at the identity of $SU(3)$, defined by
\[
\int_{SU(3)}f(V)\,\delta_I(V)\,\dd\mu_H(V)=f(I)
\]
for continuous $f$.  Define the tree-gauge delta distribution by
\[
\delta_{\mcT}(U):=\prod_{e\in\mcT}\delta_I\!\bigl(U(e)\bigr).
\]
\end{definition}

\begin{theorem}[Exact Faddeev--Popov identity in maximal-tree gauge]
\label{thm:fp-identity}
For every $U\in\mcU$,
\[
1=\int_{\mcG_*}\delta_{\mcT}(U^\Omega)\,\dd\mu_{\mcG_*}(\Omega),
\]
where $\mu_{\mcG_*}$ is product normalized Haar measure on $\mcG_*$.  Hence for
every gauge-invariant integrable functional $F$ on $\mcU$,
\[
\int_{\mcU}F(U)\,\dd\mu_{\mcU}(U)
=
\int_{\mcU}F(U)\,\delta_{\mcT}(U)\,\dd\mu_{\mcU}(U).
\]
\end{theorem}
\begin{proof}
Fix $U\in\mcU$.  By Theorem~\ref{thm:tree-gauge}, there is exactly one
$\Omega_U\in\mcG_*$ such that $U^{\Omega_U}\in\Sigma_{\mcT}$.
Therefore the integrand $\delta_{\mcT}(U^\Omega)$ vanishes for all
$\Omega\neq \Omega_U$ and contributes $1$ at $\Omega=\Omega_U$ by definition of
the delta distribution, so the integral equals $1$.

For the second identity, insert the first into the integral of $F$:
\[
\int_{\mcU}F(U)\,\dd\mu_{\mcU}(U)
=
\int_{\mcU}\int_{\mcG_*}
F(U)\,\delta_{\mcT}(U^\Omega)\,\dd\mu_{\mcG_*}(\Omega)\,\dd\mu_{\mcU}(U).
\]
Use Fubini's theorem and then change variables $U\mapsto U^{\Omega^{-1}}$.
Haar invariance implies $\dd\mu_{\mcU}(U^{\Omega^{-1}})=\dd\mu_{\mcU}(U)$, and
gauge invariance of $F$ gives $F(U^{\Omega^{-1}})=F(U)$.  Thus
\[
\int_{\mcU}F(U)\,\dd\mu_{\mcU}(U)
=
\int_{\mcU}F(U)\,\delta_{\mcT}(U)\,\dd\mu_{\mcU}(U)
\int_{\mcG_*}\dd\mu_{\mcG_*}(\Omega).
\]
The last factor equals $1$ because $\mu_{\mcG_*}$ is normalized.
\end{proof}

\paragraph{The Faddeev--Popov determinant.}
The determinant can be computed directly from infinitesimal gauge variations.
In maximal-tree gauge it is constant.

\begin{definition}[Infinitesimal gauge parameters]
\label{def:inf-gauge}
Let
\[
T^a:=\frac{\lambda_a}{2},\qquad a=1,\dots,8,
\]
where $\lambda_a$ are the Gell-Mann matrices.  For each site $x\in\mcN$ let
\[
\omega_x=\sum_{a=1}^8\omega_x^a T^a\in\mcHh,
\]
and define the infinitesimal rooted gauge transformation
\[
\Omega_\varepsilon(x):=\exp(i\varepsilon\,\omega_x),
\qquad \omega_r=0\ \text{for every }r\in R.
\]
\end{definition}

\begin{proposition}[Linearized gauge map on the tree]
\label{prop:linearized-fp}
Let $U\in\Sigma_{\mcT}$, so $U(e)=I$ for all $e\in\mcT$.  For a tree edge
$e=(x\to y)\in\mcT$,
\[
\left.\frac{\dd}{\dd\varepsilon}\right|_{\varepsilon=0}
U^{\Omega_\varepsilon}(e)
=
i(\omega_y-\omega_x).
\]
In the basis $\{T^a\}_{a=1}^8$, the linearized Faddeev--Popov operator is
\[
M_{\mathrm{FP}}^{(\mcT)} = B_{\mcT}\otimes I_8,
\]
where $B_{\mcT}$ is the rooted incidence matrix of the spanning forest.
\end{proposition}
\begin{proof}
Differentiate
\[
U^{\Omega_\varepsilon}(e)
=
\Omega_\varepsilon(y)\,U(e)\,\Omega_\varepsilon(x)^{-1}
\]
at $\varepsilon=0$.  Since $U(e)=I$ on the slice,
\[
\left.\frac{\dd}{\dd\varepsilon}\right|_{\varepsilon=0}
U^{\Omega_\varepsilon}(e)
=
i\omega_y - i\omega_x.
\]
Writing each $\omega_x$ in the Gell-Mann basis shows that the eight color
components decouple and each is acted on by the same incidence matrix
$B_{\mcT}$.
\end{proof}

\begin{proposition}[The Faddeev--Popov determinant is identically one]
\label{prop:fp-one}
With the rooted maximal-tree gauge choice,
\[
\Delta_{\mathrm{FP}}^{(\mcT)}=\det M_{\mathrm{FP}}^{(\mcT)}=1.
\]
\end{proposition}
\begin{proof}
Order the nonroot vertices so that every parent precedes every child.  Index
the rows of $B_{\mcT}$ by the tree edges $e_x=(p(x)\to x)$ and the columns by
the nonroot vertices $x$.  Then the row corresponding to $e_x$ has entry
$+1$ in column $x$, entry $-1$ in column $p(x)$ if $p(x)$ is nonroot, and zero
in all columns associated with vertices not on the path from the root to $x$.
By the chosen ordering this matrix is triangular with diagonal entries all equal
to $1$.  Hence
\[
\det B_{\mcT}=1.
\]
Therefore
\[
\det M_{\mathrm{FP}}^{(\mcT)}
=
\det(B_{\mcT}\otimes I_8)
=
(\det B_{\mcT})^8
=
1.
\]
\end{proof}

\begin{corollary}[Gauge-fixed partition function]
\label{cor:gauge-fixed-z}
The partition function may be written as
\[
Z=
\int_{\mcU\times\Lambda_{\mathrm F}\times\mcS}
\delta_{\mcT}(U)\,
e^{-S_{\mathrm W}(U)-S_{\mathrm F}(U,\bar\psi,\psi)-S_\phi(\phi)}
\dd\mu_{\mcU}(U)\,\dd\mu_{\mathrm F}\,\dd\phi.
\]
After integrating out the constrained tree links, this becomes an ordinary
integral over the cotree edge variables
\[
\mcE_{\mathrm{co}}:=\mcE\setminus\mcT.
\]
\end{corollary}
\begin{proof}
Apply Theorem~\ref{thm:fp-identity} to the gauge-invariant integrand
obtained after combining the matter variables with
Proposition~\ref{prop:gauge-measure}.  Since $\delta_{\mcT}(U)$ forces
$U(e)=I$ on all tree edges, integrating over those edges contributes the unit
factor
\[
\int_{SU(3)}\delta_I(V)\,\dd\mu_H(V)=1
\]
edge by edge.
\end{proof}

\section{Ghost fields and the effective action}
\label{sec:ghosts}

\paragraph{Introducing ghost fields.}
Although the Faddeev--Popov determinant is trivial in rooted maximal-tree
gauge, it is still useful to represent it as a finite Berezin integral, because
this makes the gauge-fixing algebra parallel to the standard continuum
formalism.

\begin{definition}[Ghost and antighost fields]
\label{def:ghosts}
For each nonroot vertex $x\in\mcN\setminus R$, let
\[
c_x=\sum_{a=1}^8 c_x^a T^a
\]
be an anticommuting $\mcHh$-valued ghost field.  For each tree edge
$e\in\mcT$, let
\[
\bar c_e=\sum_{a=1}^8 \bar c_e^a T^a
\]
be an anticommuting $\mcHh$-valued antighost field.  Their Berezin measure is
\[
\dd\mu_{\mathrm{gh}}
:=
\prod_{e\in\mcT}\prod_{a=1}^8 \dd\bar c_e^a
\prod_{x\in\mcN\setminus R}\prod_{a=1}^8 \dd c_x^a.
\]
\end{definition}

\begin{definition}[Ghost action in maximal-tree gauge]
\label{def:ghost-action}
Define
\[
S_{\mathrm{ghost}}^{(\mcT)}(\bar c,c)
:=
\sum_{e=(x\to y)\in\mcT}\tr\!\bigl(\bar c_e(c_y-c_x)\bigr),
\]
with the convention $c_r:=0$ for root vertices $r\in R$.
\end{definition}

\begin{proposition}[Ghost representation of the Faddeev--Popov determinant]
\label{prop:ghost-det}
The ghost integral reproduces the Faddeev--Popov determinant:
\[
\int e^{-S_{\mathrm{ghost}}^{(\mcT)}(\bar c,c)}\,\dd\mu_{\mathrm{gh}}
=\Delta_{\mathrm{FP}}^{(\mcT)}=1.
\]
\end{proposition}
\begin{proof}
Expand the ghost fields in the basis $\{T^a\}_{a=1}^8$.  Then the action
factorizes over color indices:
\[
S_{\mathrm{ghost}}^{(\mcT)}(\bar c,c)
=
\sum_{a=1}^8\sum_{e=(x\to y)\in\mcT}
\bar c_e^a(c_y^a-c_x^a).
\]
This is precisely the quadratic Berezin form associated with the matrix
$B_{\mcT}\otimes I_8$, so the standard finite-dimensional Berezin identity gives
\[
\int e^{-S_{\mathrm{ghost}}^{(\mcT)}}\,\dd\mu_{\mathrm{gh}}
=
\det(B_{\mcT}\otimes I_8).
\]
Proposition~\ref{prop:fp-one} gives the value $1$.
\end{proof}

\paragraph{The effective quantized action.}
To display the standard BRST structure it is convenient to introduce auxiliary
bosonic Nakanishi--Lautrup fields $b_e\in\mcHh$ for $e\in\mcT$.

\begin{definition}[BRST differential and gauge-fixing fermion]
\label{def:brst}
Let $s$ act on the full field algebra by
\[
s\,U(e)=i c_{t(e)}U(e)-iU(e)c_{s(e)},
\qquad
s\,\psi_x=i c_x\psi_x,
\qquad
s\,\bar\psi_x=-i\bar\psi_x c_x,
\]
\[
s\,\phi_x=0,
\qquad
s\,c_x=-i c_x^2,
\qquad
s\,\bar c_e=b_e,
\qquad
s\,b_e=0.
\]
Define the gauge-fixing fermion
\[
\Psi_{\mcT}
:=
\sum_{e\in\mcT}
\tr\!\left(\bar c_e\Bigl(U(e)-I+\frac{\alpha}{2}b_e\Bigr)\right),
\qquad \alpha\ge 0.
\]
\end{definition}

\begin{lemma}[Nilpotency of the BRST differential]
\label{lem:brst}
The operator $s$ satisfies $s^2=0$ on generators and therefore on the whole
field algebra.
\end{lemma}
\begin{proof}
For $\phi_x$ and $b_e$ the claim is immediate.  For $\bar c_e$,
$s^2\bar c_e=s b_e=0$.  For $c_x$,
\[
s^2 c_x = -i\,s(c_x^2) = -i\bigl((s c_x)c_x-c_x(s c_x)\bigr)
=-i\bigl((-ic_x^2)c_x-c_x(-ic_x^2)\bigr)=0.
\]
For $\psi_x$,
\[
s^2\psi_x=i(s c_x)\psi_x-i c_x(s\psi_x)
=i(-ic_x^2)\psi_x-i c_x(i c_x\psi_x)=0,
\]
and similarly for $\bar\psi_x$.  Finally, for a link variable $U(e)$,
\begin{align*}
s^2 U(e)
&=
i(s c_{t(e)})U(e)+i c_{t(e)}sU(e)-i sU(e)c_{s(e)}-iU(e)s c_{s(e)}\\
&=
c_{t(e)}^2U(e)-c_{t(e)}^2U(e)+U(e)c_{s(e)}^2-U(e)c_{s(e)}^2=0.
\end{align*}
\end{proof}

\begin{proposition}[BRST-exact gauge fixing]
\label{prop:brst-exact}
The gauge-fixing and ghost terms are
\[
S_{\mathrm{gf+gh}}^{(\mcT)}:=s\Psi_{\mcT}.
\]
Explicitly,
\[
S_{\mathrm{gf+gh}}^{(\mcT)}
=
\sum_{e\in\mcT}\tr\!\left(
b_e(U(e)-I)+\frac{\alpha}{2}b_e^2
\bar c_e\bigl(c_{t(e)}U(e)-U(e)c_{s(e)}\bigr)
\right).
\]
Consequently the effective action
\[
S_{\mathrm{eff}}
:=
S_{\mathrm W}+S_{\mathrm F}+S_\phi+S_{\mathrm{gf+gh}}^{(\mcT)}
\]
is BRST invariant:
\[
sS_{\mathrm{eff}}=0.
\]
\end{proposition}
\begin{proof}
Apply $s$ to $\Psi_{\mcT}$ and use Definition~\ref{def:brst}.  Since the
fields $b_e$ commute with everything and the ghosts anticommute, one obtains
the displayed formula.  Gauge invariance of
$S_{\mathrm W}+S_{\mathrm F}+S_\phi$ implies $s(S_{\mathrm W}+S_{\mathrm F}+S_\phi)=0$,
because $s$ is the infinitesimal gauge differential with ghost parameter $c$.
Then
\[
sS_{\mathrm{eff}}
=
s(S_{\mathrm W}+S_{\mathrm F}+S_\phi)+s^2\Psi_{\mcT}=0
\]
by Lemma~\ref{lem:brst}.
\end{proof}

\begin{remark}
At $\alpha=0$ the $b$-integration imposes the sharp gauge condition $U(e)=I$ on
tree edges.  Since the Faddeev--Popov determinant is constant, the ghost sector
decouples after gauge fixing.  The BRST presentation is therefore exact but
algebraically simpler than in continuum gauges where the determinant depends on
the gauge field.
\end{remark}

\section{Matter sector integration}
\label{sec:matter-integration}

\paragraph{Fermionic Berezin integration.}
On a finite lattice the fermion integral is exact and elementary once the
quadratic form is written in matrix notation.

\begin{definition}[Fermion matrix]
\label{def:fermion-matrix}
Order the site-color basis vectors by pairs $(x,\alpha)$ with
$x\in\mcN$ and $\alpha\in\{1,2,3\}$.  Define the matrix
$M_{\mathrm F}(U)\in\C^{3|\mcN|\times 3|\mcN|}$ by
\[
\bigl(M_{\mathrm F}(U)\bigr)_{(y,\alpha),(x,\beta)}
:=
\left(m_F+\indeg(y)\right)\delta_{xy}\delta_{\alpha\beta}
-\sum_{\substack{e\in\mcE\\ s(e)=x,\ t(e)=y}} U(e)_{\alpha\beta},
\]
where
\[
\indeg(y):=\#\{e\in\mcE:t(e)=y\}.
\]
\end{definition}

\begin{lemma}[Quadratic form identity]
\label{lem:fermion-quadratic}
With $M_{\mathrm F}(U)$ from Definition~\ref{def:fermion-matrix},
\[
S_{\mathrm F}(U,\bar\psi,\psi)
=
\sum_{x,y\in\mcN}\sum_{\alpha,\beta=1}^3
\bar\psi_{y,\alpha}\,
\bigl(M_{\mathrm F}(U)\bigr)_{(y,\alpha),(x,\beta)}\,
\psi_{x,\beta}.
\]
\end{lemma}
\begin{proof}
Expand the edge sum in Definition~\ref{def:imported-package}:
\[
\sum_{e\in\mcE}\bar\psi_{t(e)}\psi_{t(e)}
=
\sum_{y\in\mcN}\indeg(y)\,\bar\psi_y\psi_y.
\]
The remaining transported terms contribute the off-diagonal blocks
$-\sum_{e:s(e)=x,t(e)=y}U(e)$.  Adding the mass term gives the stated matrix.
\end{proof}

\begin{theorem}[Fermion determinant]
\label{thm:fermion-det}
For every gauge field $U\in\mcU$,
\[
\int_{\Lambda_{\mathrm F}}
e^{-S_{\mathrm F}(U,\bar\psi,\psi)}\,\dd\mu_{\mathrm F}
=
\det M_{\mathrm F}(U).
\]
Consequently
\[
Z
=
\int_{\mcU\times\mcS}
e^{-S_{\mathrm W}(U)-S_\phi(\phi)}\det M_{\mathrm F}(U)\,
\dd\mu_{\mcU}(U)\,\dd\phi.
\]
\end{theorem}
\begin{proof}
By Lemma~\ref{lem:fermion-quadratic}, the exponent is the standard quadratic
Berezin form $\bar\Psi M_{\mathrm F}(U)\Psi$ in finitely many Grassmann
variables.  The finite-dimensional Berezin integral identity therefore gives
\[
\int e^{-\bar\Psi M\Psi}\,\dd\bar\Psi\,\dd\Psi=\det M
\]
for any finite matrix $M$.  Applying this with $M=M_{\mathrm F}(U)$ yields the
first claim.  Substituting the result into the definition of $Z$ gives the
second claim.
\end{proof}

\begin{remark}[Virtual loops and graph support]
\label{rem:fermion-loops}
The determinant $\det M_{\mathrm F}(U)$ is a finite polynomial in the link
matrices and encodes the usual closed virtual-fermion loops.  Because the link
set $\mcE$ consists of the worldline and interaction edges from the trajectory
data, only loops in that explicit graph contribute.  The interaction-graph
weights $\omega_{ij}^{\mathrm{link}}(t)$ from \cite{latticeqft} influence these
virtual loops through the dependence of the original data-induced background
links on the viscous force and its weights.
\end{remark}

\paragraph{Scalar kinetic fluctuations.}
The scalar action is purely bosonic, real valued, and bounded below as soon as
the local potential is bounded below.  The regularization $\eta>0$ excludes
edge-length singularities.

\begin{definition}[Scalar stiffness operator]
\label{def:scalar-operator}
Define the symmetric quadratic form
\[
Q_\eta(\phi)
:=
\frac12\sum_{e=(x\to y)\in\mcE}\frac{(\phi_y-\phi_x)^2}{\ell_e+\eta}.
\]
There is a unique symmetric matrix $L_\eta\in\R^{|\mcN|\times|\mcN|}$ such that
\[
Q_\eta(\phi)=\frac12\,\phi^\top L_\eta \phi.
\]
\end{definition}

\begin{lemma}[Positivity and bounded coefficients]
\label{lem:scalar-positive}
The operator $L_\eta$ is positive semidefinite.  Moreover every edge weight
\[
w_e:=\frac{1}{\ell_e+\eta}
\]
satisfies
\[
0<w_e\le \eta^{-1}.
\]
\end{lemma}
\begin{proof}
Each summand in $Q_\eta(\phi)$ is nonnegative, so $Q_\eta(\phi)\ge 0$ for every
$\phi$, hence $L_\eta$ is positive semidefinite.  Since $\ell_e\ge 0$ and
$\eta>0$,
\[
\ell_e+\eta\ge \eta,
\]
which implies the upper bound on $w_e$.
\end{proof}

\begin{theorem}[Scalar integrability]
\label{thm:scalar-integrable}
Assume $V_\phi$ is continuous and bounded below by $-C_\phi$.  Then
\[
Z_\phi
:=
\int_{\mcS}e^{-S_\phi(\phi)}\,\dd\phi
\]
is finite.  If in addition $V_\phi\equiv 0$, then
\[
Z_\phi
=
(2\pi)^{|\mcN|/2}
\det(L_\eta+m_\phi^2 I)^{-1/2}.
\]
\end{theorem}
\begin{proof}
By Definition~\ref{def:scalar-operator},
\[
S_\phi(\phi)=\frac12\phi^\top L_\eta\phi
+\frac{m_\phi^2}{2}\phi^\top\phi
+\sum_{x\in\mcN}V_\phi(\phi_x).
\]
Lemma~\ref{lem:scalar-positive} implies the first quadratic term is
nonnegative, while the lower bound on $V_\phi$ gives
\[
S_\phi(\phi)\ge \frac{m_\phi^2}{2}\phi^\top\phi-C_\phi|\mcN|.
\]
Thus
\[
e^{-S_\phi(\phi)}
\le
e^{C_\phi|\mcN|}e^{-\frac12 m_\phi^2 \phi^\top\phi},
\]
and the right-hand side is integrable on $\R^{|\mcN|}$.

If $V_\phi\equiv 0$, then
\[
S_\phi(\phi)=\frac12\phi^\top (L_\eta+m_\phi^2 I)\phi.
\]
Since $L_\eta$ is positive semidefinite and $m_\phi^2 I$ is strictly positive,
the matrix $L_\eta+m_\phi^2 I$ is positive definite.  The standard Gaussian
integral formula gives the displayed determinant expression.
\end{proof}

\begin{remark}[Role of $\eta$]
\label{rem:eta}
The parameter $\eta>0$ is not a perturbative convenience; it is the exact
mechanism that prevents the scalar kinetic coefficients from diverging when two
sites become spacetime-coincident.  Without $\eta$, the coefficient
$1/\ell_e$ would be singular on zero-length edges.  With $\eta>0$, every
kinetic coefficient is uniformly bounded by $\eta^{-1}$.
\end{remark}

\section{Discussion of excitations and propagators}
\label{sec:propagators}

\paragraph{The propagator on the interaction graph.}
Because the theory is finite, all correlation functions can be defined by
source differentiation exactly.  Their graph-theoretic support is determined by
the explicit trajectory-induced lattice.

\begin{definition}[Generating functional]
\label{def:generating}
Introduce scalar sources $J=(J_x)_{x\in\mcN}$ and fermionic sources
$\eta=(\eta_x)_{x\in\mcN}$, $\bar\eta=(\bar\eta_x)_{x\in\mcN}$.  The generating
functional is
\[
Z[J,\eta,\bar\eta]
:=
\int
\exp\!\left(
-S
+\sum_{x\in\mcN}J_x\phi_x
+\sum_{x\in\mcN}\bar\eta_x\psi_x
+\sum_{x\in\mcN}\bar\psi_x\eta_x
\right)\dd\mu.
\]
\end{definition}

\begin{definition}[Two-point functions]
\label{def:twopoint}
Whenever $Z[0,0,0]\neq 0$, the scalar and fermion two-point functions are
\[
G_\phi(x,y):=\langle \phi_x\phi_y\rangle,
\qquad
S_F(x,y)_{\alpha\beta}:=\langle \psi_{x,\alpha}\bar\psi_{y,\beta}\rangle.
\]
\end{definition}

\begin{proposition}[Background-field propagators]
\label{prop:background-prop}
Fix a gauge field $U$.
\begin{enumerate}[leftmargin=1.5em]
\item If $V_\phi\equiv 0$, then the scalar background covariance is
\[
G_\phi^{(U)}=(L_\eta+m_\phi^2 I)^{-1}.
\]
\item If $M_{\mathrm F}(U)$ is invertible, then the fermion background
covariance is
\[
S_F^{(U)}=M_{\mathrm F}(U)^{-1}.
\]
\end{enumerate}
\end{proposition}
\begin{proof}
The scalar claim is the standard Gaussian covariance formula from
Theorem~\ref{thm:scalar-integrable}.  The fermion claim follows from
differentiating the Berezin generating functional
\[
\int e^{-\bar\Psi M_{\mathrm F}(U)\Psi+\bar\eta\Psi+\bar\Psi\eta}\,\dd\bar\Psi\,\dd\Psi
=
\det M_{\mathrm F}(U)\,
e^{\bar\eta M_{\mathrm F}(U)^{-1}\eta}.
\]
\end{proof}

\begin{proposition}[Large-mass hopping expansion]
\label{prop:hopping}
Assume $M_{\mathrm F}(U)=D-H(U)$ where
\[
D:=\diag\bigl(m_F+\indeg(x)\bigr)_{x\in\mcN}\otimes I_3
\]
and suppose
\[
\|D^{-1}H(U)\|<1
\]
in operator norm.  Then
\[
M_{\mathrm F}(U)^{-1}
=
\sum_{n=0}^\infty (D^{-1}H(U))^n D^{-1},
\]
and the $(x,y)$ matrix element is a convergent sum over directed paths
$\gamma:y\rightsquigarrow x$ in $\mcE$.  Therefore the support of the fermion
propagator is controlled by the explicit interaction/worldline graph, and the
same-time path combinatorics are inherited from the interaction data
$\mathcal P_t$ and hence from the stochastic weights
$\omega_{ij}^{\mathrm{link}}(t)$ used in \cite{latticeqft}.
\end{proposition}
\begin{proof}
The Neumann series for $(I-D^{-1}H)^{-1}$ converges in operator norm under the
assumed bound.  Each power $(D^{-1}H)^n$ is a sum over length-$n$ compositions
of off-diagonal hopping operators, hence over directed paths of length $n$ in
the graph.  The only same-time edges available for those paths are the
interaction edges determined by the data sets $\mathcal P_t$, which were
constructed from the weights $\omega_{ij}^{\mathrm{link}}(t)$.
\end{proof}

\paragraph{Wilson loops and the area law.}
The Wilson loop remains the basic gauge-invariant probe of confinement.

\begin{definition}[Wilson loop expectation]
\label{def:wilson-loop-exp}
For an oriented closed loop $\gamma$ in the lattice,
\[
W(\gamma):=\tr\!\left(\prod_{e\subset \gamma}U(e)\right),
\qquad
\langle W(\gamma)\rangle := Z^{-1}\int W(\gamma)e^{-S}\,\dd\mu,
\]
whenever $Z\neq 0$.
\end{definition}

\begin{remark}[Area-law criterion]
\label{rem:area-law}
If a family of loops $\gamma$ admits spanning surfaces $\Sigma_\gamma$ in the
embedded spacetime geometry and one can prove
\[
|\langle W(\gamma)\rangle|
\le
C\,e^{-\sigma\,\mathrm{Area}(\Sigma_\gamma)}
\]
with $\sigma>0$ independent of the loop size, then one has the usual Euclidean
area-law signature of confinement \cite{wilson1974confinement}.  The present
paper does \emph{not} establish such a bound; it merely provides the exact
finite-lattice measure in which the question is well posed.
\end{remark}

\section{Conclusion}
\label{sec:conclusion}

\paragraph{Summary of the quantum data package.}
Starting from the finite lattice data package $\mathcal Q$ built in
\cite{latticeqft}, this paper has completed the exact finite-dimensional
quantization step.  The link variables were promoted from data-derived
transporters to independent $SU(3)$ variables, the full configuration space and
product measure were defined, and the partition function was proved finite under
the minimal scalar lower-bound assumption needed for scalar integrability.  The
gauge redundancy was solved globally by a rooted maximal-tree gauge, with exact
orbit uniqueness, exact Faddeev--Popov identity, constant determinant, and a
decoupled ghost sector.  The fermions integrate to a finite determinant, the
scalar sector is controlled by a positive stiffness operator regularized by
$\eta$, and the bilocal mean-field action $S_{\mathrm W}^\infty$ defined on the
stationary measure $\pi_\infty$ provides the correct thermodynamic-limit
continuum theory.

\paragraph{Forward look to the mass-gap problem.}
The $N\to\infty$ thermodynamic limit provides the correct continuum theory:
the bilocal action $S_{\mathrm W}^\infty$, the continuous propagator as the
resolvent of $K_\rho(z,z')$, and the Markov-chain measure converging to a
continuous process on $\pi_\infty$.
The remaining steps toward the Millennium problem are: verification of the
Osterwalder--Schrader axioms on the integral-operator continuum theory;
reconstruction of a Hilbert-space quantum theory from reflection positivity;
and a proof that the energy--momentum spectrum has a strictly positive gap
above the vacuum.
What this paper contributes is the exact finite-dimensional quantization step
on the trajectory-induced lattice, together with the bilocal mean-field
structure that survives the $N\to\infty$ limit.

\bibliographystyle{plain}
\bibliography{references}

\end{document}
